### 2015-02-21 manuscript/manuscript.tex

% Template for PLoS
% Version 3.1 February 2015
%
% To compile to pdf, run:
% latex plos.template
% bibtex plos.template
% latex plos.template
% latex plos.template
% dvipdf plos.template
%
% % % % % % % % % % % % % % % % % % % % % %
%
% -- IMPORTANT NOTE
%
% This template contains comments intended 
% to minimize problems and delays during our production 
% process. Please follow the template instructions
% whenever possible.
%
% % % % % % % % % % % % % % % % % % % % % % % 
%
% Once your paper is accepted for publication, 
% PLEASE REMOVE ALL TRACKED CHANGES in this file and leave only
% the final text of your manuscript.
%
% There are no restrictions on package use within the LaTeX files except that 
% no packages listed in the template may be deleted.
%
% Please do not include colors or graphics in the text.
%
% Please do not create a heading level below \subsection. For 3rd level headings, use \paragraph{}.
%
% % % % % % % % % % % % % % % % % % % % % % %
%
% -- FIGURES AND TABLES
%
% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
%
% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
% - Figures should be uploaded separately from your manuscript file. 
% - Figures generated using LaTeX should be extracted and removed from the PDF before submission. 
% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
% For figure citations, please use "Fig." instead of "Figure".
% See http://www.plosone.org/static/figureGuidelines for PLOS figure guidelines.
%
% Tables should be cell-based and may not contain:
% - tabs/spacing/line breaks within cells to alter layout or alignment
% - vertically-merged cells (no tabular environments within tabular environments, do not use \multirow)
% - colors, shading, or graphic objects
% See http://www.plosone.org/static/figureGuidelines#tables for table guidelines.
%
% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
%
% IMPORTANT
% Below are a few tips to help format your equations and other special characters according to our specifications. For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://www.plosone.org/static/latexGuidelines
%
% Please be sure to include all portions of an equation in the math environment.
%
% Do not include text that is not math in the math environment. For example, CO2 will be CO\textsubscript{2}.
%
% Please add line breaks to long display equations when possible in order to fit size of the column. 
%
% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
%
% % % % % % % % % % % % % % % % % % % % % % % % 
%
% Please contact latex@plos.org with any questions.
%
% % % % % % % % % % % % % % % % % % % % % % % %

\documentclass[10pt,letterpaper]{article}
\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}

% Use adjustwidth environment to exceed column width (see example table in text)
\usepackage{changepage}

% Use Unicode characters when possible
\usepackage[utf8]{inputenc}

% textcomp package and marvosym package for additional characters
\usepackage{textcomp,marvosym}

% fixltx2e package for \textsubscript
\usepackage{fixltx2e}

% amsmath and amssymb packages, useful for mathematical formulas and symbols
\usepackage{amsmath,amssymb}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

% Use nameref to cite supporting information files (see Supporting Information section for more info)
\usepackage{nameref,hyperref}

% line numbers
\usepackage[right]{lineno}

% ligatures disabled
\usepackage{microtype}
\DisableLigatures[f]{encoding = *, family = * }

% rotating package for sideways tables
\usepackage{rotating}

% Remove comment for double spacing
%\usepackage{setspace} 
%\doublespacing

% Text layout
\raggedright
\setlength{\parindent}{0.5cm}
\textwidth 5.25in 
\textheight 8.75in

% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
% Captions will be left justified
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}

% Use the PLoS provided BiBTeX style
\bibliographystyle{resources/plos2015}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Leave date blank
\date{}

% Header and Footer with logo
\usepackage{lastpage,fancyhdr,graphicx}
\usepackage{epstopdf}
\pagestyle{myheadings}
\pagestyle{fancy}
\fancyhf{}
\lhead{\includegraphics[width=2.0in]{resources/PLOS-Submission.eps}}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
\fancyheadoffset[L]{2.25in}
\fancyfootoffset[L]{2.25in}
\lfoot{\sf PLOS}

%% Include all macros below

\newcommand{\lorem}{{\bf LOREM}}
\newcommand{\ipsum}{{\bf IPSUM}}

%% END MACROS SECTION

\begin{document}
\vspace*{0.35in}

% Title must be 250 characters or less.
% Please capitalize all terms in the title except conjunctions, prepositions, and articles.

\begin{flushleft}
{\Large
\textbf{Ten Simple Rules for Data Storage}
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Edmund Hart $^{1 ,\ast}$,
Pauline Barmby $^{2 }$,
Jeffrey Hollister $^{3 }$,
\\
\bigskip
\bf{1} Univeristy of Vermont, Department of Biology, Burlington
\\
\bf{2} University of Western Ontario, Department of Physics and Astronomy
\\
\bf{3} US Environmental Protection Agency, Atlantic Ecology Division
\\
\bigskip
 $\ast$ E-mail: emh@emhart.info

\end{flushleft}

\section*{Abstract}\label{abstract}
\addcontentsline{toc}{section}{Abstract}

\linenumbers

\section*{Introduction}\label{introduction}
\addcontentsline{toc}{section}{Introduction}

Some example text with a citation {[}1{]}

\section*{Rule 1: Rule}\label{rule-1-rule}
\addcontentsline{toc}{section}{Rule 1: Rule}

\section*{Rule 2: Rule}\label{rule-2-rule}
\addcontentsline{toc}{section}{Rule 2: Rule}

\section*{Rule 3: Rule}\label{rule-3-rule}
\addcontentsline{toc}{section}{Rule 3: Rule}

\section*{Rule 4: Rule}\label{rule-4-rule}
\addcontentsline{toc}{section}{Rule 4: Rule}

\section*{Rule 5: Rule}\label{rule-5-rule}
\addcontentsline{toc}{section}{Rule 5: Rule}

\section*{Rule 6: Rule}\label{rule-6-rule}
\addcontentsline{toc}{section}{Rule 6: Rule}

\section*{Rule 7: Rule}\label{rule-7-rule}
\addcontentsline{toc}{section}{Rule 7: Rule}

\section*{Rule 8: Rule}\label{rule-8-rule}
\addcontentsline{toc}{section}{Rule 8: Rule}

\section*{Rule 9: Rule}\label{rule-9-rule}
\addcontentsline{toc}{section}{Rule 9: Rule}

\section*{Rule 10: Rule}\label{rule-10-rule}
\addcontentsline{toc}{section}{Rule 10: Rule}

\section*{Figure Legends}\label{figure-legends}
\addcontentsline{toc}{section}{Figure Legends}

Figures here: Will need to figure out numbering\ldots{}

\section*{Tables}\label{tables}
\addcontentsline{toc}{section}{Tables}

Tables here: Will need to figure out numbering\ldots{}

\nolinenumbers

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\bibdata{resources/manuscript.bib}

1. Goodman A, Pepe A, Blocker AW, Borgman CL, Cranmer K, Crosas M, et
al. Ten simple rules for the care and feeding of scientific data. PLoS
computational biology. Public Library of Science; 2014;10: e1003542.

% Please keep the abstract below 300 words

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLOS ONE authors please skip this step. 
% Author Summary not valid for PLOS ONE submissions.

% For figure citations, please use "Fig." instead of "Figure".

%\section*{References}
% Either type in your references using
% \begin{thebibliography}{}
% \bibitem{}
% Text
% \end{thebibliography}
%
% OR
%
% Compile your BiBTeX database using our plos2015.bst
% style file and paste the contents of your .bbl file
% here.
%

\end{document}

### 2015-02-21 manuscript/resources/plos_template_orig.tex

--- 2015-02-21 manuscript/manuscript.tex
@@
 % Use the PLoS provided BiBTeX style
-\bibliographystyle{resources/plos2015}
+\bibliographystyle{plos2015}
 
@@
 \fancyhf{}
-\lhead{\includegraphics[width=2.0in]{resources/PLOS-Submission.eps}}
+\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
 \rfoot{\thepage/\pageref{LastPage}}
@@
 % Please capitalize all terms in the title except conjunctions, prepositions, and articles.
-
 \begin{flushleft}
 {\Large
-\textbf{Ten Simple Rules for Data Storage}
+\textbf\newline{Article Title - Submission to PLOS Journals}
 }
@@
 \\
-Edmund Hart $^{1 ,\ast}$,
-Pauline Barmby $^{2 }$,
-Jeffrey Hollister $^{3 }$,
+Name1 Surname\textsuperscript{1,\Yinyang},
+Name2 Surname\textsuperscript{2,\Yinyang},
+Name3 Surname\textsuperscript{2,\textcurrency a},
+Name4 Surname\textsuperscript{2,\ddag},
+Name5 Surname\textsuperscript{2,\ddag},
+Name6 Surname\textsuperscript{2},
+Name7 Surname\textsuperscript{3,*},
+with the Lorem Ipsum Consortium\textsuperscript{\textpilcrow}
 \\
 \bigskip
-\bf{1} Univeristy of Vermont, Department of Biology, Burlington
+\bf{1} Affiliation Dept/Program/Center, Institution Name, City, State, Country
 \\
-\bf{2} University of Western Ontario, Department of Physics and Astronomy
+\bf{2} Affiliation Dept/Program/Center, Institution Name, City, State, Country
 \\
-\bf{3} US Environmental Protection Agency, Atlantic Ecology Division
+\bf{3} Affiliation Dept/Program/Center, Institution Name, City, State, Country
 \\
 \bigskip
- $\ast$ E-mail: emh@emhart.info
+
+% Insert additional author notes using the symbols described below. Insert symbol callouts after author names as necessary.
+% 
+% Remove or comment out the author notes below if they aren't used.
+%
+% Primary Equal Contribution Note
+\Yinyang These authors contributed equally to this work.
+
+% Additional Equal Contribution Note
+% Also use this double-dagger symbol for special authorship notes, such as senior authorship.
+\ddag These authors also contributed equally to this work.
+
+% Current address notes
+\textcurrency a Insert current address of first author with an address update
+% \textcurrency b Insert current address of second author with an address update
+% \textcurrency c Insert current address of third author with an address update
+
+% Deceased author note
+\dag Deceased
+
+% Group/Consortium Author Note
+\textpilcrow Membership list can be found in the Acknowledgments section.
+
+% Use the asterisk to denote corresponding authorship and provide email address in note below.
+* CorrespondingAuthor@institute.edu
 
 \end{flushleft}
-
-\section*{Abstract}\label{abstract}
-\addcontentsline{toc}{section}{Abstract}
-
-\linenumbers
-
-\section*{Introduction}\label{introduction}
-\addcontentsline{toc}{section}{Introduction}
-
-Some example text with a citation {[}1{]}
-
-\section*{Rule 1: Rule}\label{rule-1-rule}
-\addcontentsline{toc}{section}{Rule 1: Rule}
-
-\section*{Rule 2: Rule}\label{rule-2-rule}
-\addcontentsline{toc}{section}{Rule 2: Rule}
-
-\section*{Rule 3: Rule}\label{rule-3-rule}
-\addcontentsline{toc}{section}{Rule 3: Rule}
-
-\section*{Rule 4: Rule}\label{rule-4-rule}
-\addcontentsline{toc}{section}{Rule 4: Rule}
-
-\section*{Rule 5: Rule}\label{rule-5-rule}
-\addcontentsline{toc}{section}{Rule 5: Rule}
-
-\section*{Rule 6: Rule}\label{rule-6-rule}
-\addcontentsline{toc}{section}{Rule 6: Rule}
-
-\section*{Rule 7: Rule}\label{rule-7-rule}
-\addcontentsline{toc}{section}{Rule 7: Rule}
-
-\section*{Rule 8: Rule}\label{rule-8-rule}
-\addcontentsline{toc}{section}{Rule 8: Rule}
-
-\section*{Rule 9: Rule}\label{rule-9-rule}
-\addcontentsline{toc}{section}{Rule 9: Rule}
-
-\section*{Rule 10: Rule}\label{rule-10-rule}
-\addcontentsline{toc}{section}{Rule 10: Rule}
-
-\section*{Figure Legends}\label{figure-legends}
-\addcontentsline{toc}{section}{Figure Legends}
-
-Figures here: Will need to figure out numbering\ldots{}
-
-\section*{Tables}\label{tables}
-\addcontentsline{toc}{section}{Tables}
-
-Tables here: Will need to figure out numbering\ldots{}
-
-\nolinenumbers
-
-\section*{References}\label{references}
-\addcontentsline{toc}{section}{References}
-
-\bibdata{resources/manuscript.bib}
-
-1. Goodman A, Pepe A, Blocker AW, Borgman CL, Cranmer K, Crosas M, et
-al. Ten simple rules for the care and feeding of scientific data. PLoS
-computational biology. Public Library of Science; 2014;10: e1003542.
-
 % Please keep the abstract below 300 words
+\section*{Abstract}
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Curabitur eget porta erat.
+Morbi consectetur est vel gravida pretium.
+Suspendisse ut dui eu ante cursus gravida non sed sem.
+Nullam sapien tellus, commodo id velit id, eleifend volutpat quam.
+Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus.
+Nunc pellentesque pretium diam, quis maximus dolor faucibus id.
+Nunc convallis sodales ante, ut ullamcorper est egestas vitae.
+Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.
 
@@
 % Author Summary not valid for PLOS ONE submissions.   
+\section*{Author Summary}
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Curabitur eget porta erat.
+Morbi consectetur est vel gravida pretium.
+Suspendisse ut dui eu ante cursus gravida non sed sem.
+Nullam sapien tellus, commodo id velit id, eleifend volutpat quam.
+Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus.
+Nunc pellentesque pretium diam, quis maximus dolor faucibus id.
+Nunc convallis sodales ante, ut ullamcorper est egestas vitae.
+Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.
+
+\linenumbers
+
+\section*{Introduction}
+Lorem ipsum dolor sit~\cite{bib1} amet, consectetur adipiscing elit.
+Curabitur eget porta erat.
+Morbi consectetur est vel gravida pretium.
+Suspendisse ut dui eu ante cursus gravida non sed sem.
+Nullam Eq.~(\ref{eq:schemeP}) sapien tellus, commodo id velit id, eleifend volutpat quam.
+Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus.
+Nunc pellentesque pretium diam, quis maximus dolor faucibus id.~\cite{bib2} Nunc convallis sodales ante, ut ullamcorper est egestas vitae.
+Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.
+
+\begin{equation}\label{eq:schemeP} 
+D_{coll} = \frac{D_f+\frac{[S]^2}{K_D S_T} D_S} {1+\frac{[S]^2}{K_D S_T}}, 
+D_{sm} = \frac{D_f+ \frac{[S]}{K_D} D_S}{1+\frac{[S]}{K_D}},
+\end{equation}
+
+% You may title this section "Methods" or "Models". 
+% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
+% authors may use "Analysis" 
+\section*{Materials and Methods}
+\subsection*{Etiam eget sapien nibh.}
 
 % For figure citations, please use "Fig." instead of "Figure".
+Nulla mi mi, Fig.~\ref{fig1} venenatis sed ipsum varius, volutpat euismod diam.
+Proin rutrum vel massa non gravida.
+Quisque tempor sem et dignissim rutrum.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Morbi at justo vitae nulla elementum commodo eu id massa.
+In vitae diam ac augue semper tincidunt eu ut eros.
+Fusce fringilla erat porttitor lectus cursus, \nameref{S1_Video} vel sagittis arcu lobortis.
+Aliquam in enim semper, aliquam massa id, cursus neque.
+Praesent faucibus semper libero.
+
+\begin{figure}[h]
+\caption{{\bf Figure Title first bold sentence Nulla mi mi, venenatis sed ipsum varius, volutpat euismod diam.}
+Figure Caption Proin rutrum vel massa non gravida. Quisque tempor sem et dignissim rutrum. A: Lorem ipsum dolor sit amet. B: Consectetur adipiscing elit.}
+\label{fig1}
+\end{figure}
+
+\begin{enumerate}
+\item{react}
+\item{diffuse free particles}
+\item{increment time by dt and go to 1}
+\end{enumerate}
+
+% Results and Discussion can be combined.
+\section*{Results}
+Nulla mi mi, venenatis sed ipsum varius, Table~\ref{table1} volutpat euismod diam.
+Proin rutrum vel massa non gravida.
+Quisque tempor sem et dignissim rutrum.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Morbi at justo vitae nulla elementum commodo eu id massa.
+In vitae diam ac augue semper tincidunt eu ut eros.
+Fusce fringilla erat porttitor lectus cursus, vel sagittis arcu lobortis.
+Aliquam in enim semper, aliquam massa id, cursus neque.
+Praesent faucibus semper libero.
+
+\begin{table}[!ht]
+\begin{adjustwidth}{-2.25in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
+\caption{
+{\bf Table caption Nulla mi mi, venenatis sed ipsum varius, volutpat euismod diam.}}
+\begin{tabular}{|l|l|l|l|l|l|l|l|}
+\hline
+\multicolumn{4}{|l|}{\bf Heading1} & \multicolumn{4}{|l|}{\bf Heading2}\\ \hline
+$cell1 row1$ & cell2 row 1 & cell3 row 1 & cell4 row 1 & cell5 row 1 & cell6 row 1 & cell7 row 1 & cell8 row 1\\ \hline
+$cell1 row2$ & cell2 row 2 & cell3 row 2 & cell4 row 2 & cell5 row 2 & cell6 row 2 & cell7 row 2 & cell8 row 2\\ \hline
+$cell1 row3$ & cell2 row 3 & cell3 row 3 & cell4 row 3 & cell5 row 3 & cell6 row 3 & cell7 row 3 & cell8 row 3\\ \hline
+\end{tabular}
+\begin{flushleft} Table notes Phasellus venenatis, tortor nec vestibulum mattis, massa tortor interdum felis, nec pellentesque metus tortor nec nisl.
+Ut ornare mauris tellus, vel dapibus arcu suscipit sed.
+\end{flushleft}
+\label{table1}
+\end{adjustwidth}
+\end{table}
+
+\subsection*{\lorem\ and \ipsum\ Nunc blandit a tortor.}
+
+Maecenas convallis mauris sit amet sem ultrices gravida.
+Etiam eget sapien nibh.
+Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula.
+Curabitur fringilla pulvinar lectus consectetur pellentesque.
+Quisque augue sem, tincidunt sit amet feugiat eget, ullamcorper sed velit.
+Sed non aliquet felis.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Mauris commodo justo ac dui pretium imperdiet.
+Sed suscipit iaculis mi at feugiat.
+
+\subsection*{Sed ac quam id nisi malesuada congue.}
+
+Nulla mi mi, venenatis sed ipsum varius, volutpat euismod diam.
+Proin rutrum vel massa non gravida.
+Quisque tempor sem et dignissim rutrum.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Morbi at justo vitae nulla elementum commodo eu id massa.
+In vitae diam ac augue semper tincidunt eu ut eros.
+Fusce fringilla erat porttitor lectus cursus, vel sagittis arcu lobortis.
+Aliquam in enim semper, aliquam massa id, cursus neque.
+Praesent faucibus semper libero.
+
+% Please do not create a heading level below \subsection. For 3rd level headings, use \paragraph{}. 
+\subsection*{Subsection 1}
+Nulla mi mi, venenatis sed ipsum varius, volutpat euismod diam.
+Proin rutrum vel massa non gravida.
+Quisque tempor sem et dignissim rutrum.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Morbi at justo vitae nulla elementum commodo eu id massa.
+In vitae diam ac augue semper tincidunt eu ut eros.
+Fusce fringilla erat porttitor lectus cursus, vel sagittis arcu lobortis.
+Aliquam in enim semper, aliquam massa id, cursus neque.
+Praesent faucibus semper libero.
+
+\subsection*{Subsection 2}
+\paragraph{3rd Level Heading.} Nulla mi mi, venenatis sed ipsum varius, volutpat euismod diam.
+Proin rutrum vel massa non gravida.
+Quisque tempor sem et dignissim rutrum.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Morbi at justo vitae nulla elementum commodo eu id massa.
+In vitae diam ac augue semper tincidunt eu ut eros.
+Fusce fringilla erat porttitor lectus cursus, vel sagittis arcu lobortis.
+Aliquam in enim semper, aliquam massa id, cursus neque.
+Praesent faucibus semper libero.
+
+\section*{Discussion}
+Nulla mi mi, venenatis sed ipsum varius, Table~\ref{table1} volutpat euismod diam.
+Proin rutrum vel massa non gravida.
+Quisque tempor sem et dignissim rutrum.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Morbi at justo vitae nulla elementum commodo eu id massa.
+In vitae diam ac augue semper tincidunt eu ut eros.
+Fusce fringilla erat porttitor lectus cursus, vel sagittis arcu lobortis.
+Aliquam in enim semper, aliquam massa id, cursus neque.
+Praesent faucibus semper libero.
+
+\subsection*{\lorem\ and \ipsum\ Nunc blandit a tortor.}
+
+CO\textsubscript{2} Maecenas convallis mauris sit amet sem ultrices gravida.
+Etiam eget sapien nibh.
+Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula.
+Curabitur fringilla pulvinar lectus consectetur pellentesque.
+Quisque augue sem, tincidunt sit amet feugiat eget, ullamcorper sed velit.
+
+Sed non aliquet felis.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Mauris commodo justo ac dui pretium imperdiet.
+Sed suscipit iaculis mi at feugiat.
+Ut neque ipsum, luctus id lacus ut, laoreet scelerisque urna.
+Phasellus venenatis, tortor nec vestibulum mattis, massa tortor interdum felis, nec pellentesque metus tortor nec nisl.
+Ut ornare mauris tellus, vel dapibus arcu suscipit sed.
+Nam condimentum sem eget mollis euismod.
+Nullam dui urna, gravida venenatis dui et, tincidunt sodales ex.
+Nunc est dui, sodales sed mauris nec, auctor sagittis leo.
+Aliquam tincidunt, ex in facilisis elementum, libero lectus luctus est, non vulputate nisl augue at dolor.
+For more information, see \nameref{S1_Text}.
+
+\section*{Supporting Information}
+
+% Include only the SI item label in the subsection heading. Use the \nameref{label} command to cite SI items in the text.
+\subsection*{S1 Video}
+\label{S1_Video}
+{\bf Bold the first sentence.}  Maecenas convallis mauris sit amet sem ultrices gravida.
+Etiam eget sapien nibh.
+Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula.
+Curabitur fringilla pulvinar lectus consectetur pellentesque.
+
+\subsection*{S1 Text}
+\label{S1_Text}
+{\bf Lorem Ipsum.} Maecenas convallis mauris sit amet sem ultrices gravida.
+Etiam eget sapien nibh.
+Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula.
+Curabitur fringilla pulvinar lectus consectetur pellentesque.
+
+\subsection*{S1 Fig}
+\label{S1_Fig}
+{\bf Lorem Ipsum.} Maecenas convallis mauris sit amet sem ultrices gravida.
+Etiam eget sapien nibh.
+Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula.
+Curabitur fringilla pulvinar lectus consectetur pellentesque.
+
+\subsection*{S2 Fig}
+\label{S2_Fig}
+{\bf Lorem Ipsum.} Maecenas convallis mauris sit amet sem ultrices gravida.
+Etiam eget sapien nibh.
+Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula.
+Curabitur fringilla pulvinar lectus consectetur pellentesque.
+
+\subsection*{S1 Table}
+\label{S1_Table}
+{\bf Lorem Ipsum.} Maecenas convallis mauris sit amet sem ultrices gravida.
+Etiam eget sapien nibh.
+Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula.
+Curabitur fringilla pulvinar lectus consectetur pellentesque.
+
+\section*{Acknowledgments}
+Cras egestas velit mauris, eu mollis turpis pellentesque sit amet.
+Interdum et malesuada fames ac ante ipsum primis in faucibus.
+Nam id pretium nisi.
+Sed ac quam id nisi malesuada congue.
+Sed interdum aliquet augue, at pellentesque quam rhoncus vitae.
+
+\nolinenumbers
 
@@
 % 
+\begin{thebibliography}{10}
+\bibitem{bib1}
+Devaraju P, Gulati R, Antony PT, Mithun CB, Negi VS.
+Susceptibility to SLE in South Indian Tamils may be influenced by genetic selection pressure on TLR2 and TLR9 genes.
+Mol Immunol.
+2014 Nov 22.
+pii: S0161-5890(14)00313-7.
+doi: 10.1016/j.molimm.2014.11.005
+
+\bibitem{bib2}
+Huynen MMTE, Martens P, Hilderlink HBM. The health impacts of globalisation: a conceptual framework. Global Health. 2005;1: 14. Available: http://www.globalizationandhealth.com/content/1/1/14.
+
+\end{thebibliography}
 

### 2016-06-01 resubmission_materials/reviewer_response.md
added a response letter

Dear Dr. Markel,

Thank you for reviewing our manuscript "Ten Simple Rules for Digital Data Storage" and considering our resubmission. We appreciate the thoughtful feedback reviewers have provided and incorporated their comments into our revised manuscript. Below is a detailed accounting of our changes, with reviewer submissions highlighted in bold.

**_Reviewer #1:_** _The manuscript enumerates a set of "rules" that promote up-front thought during experimental design regarding the long term storage of data generated. Each of the ten rules has a specific focus on one aspect that researchers should consider. The authors clearly communicate the rules with relevant supporting observations.
A few minor issues:_

**Rules 4 and 5 seem like variations on the same rule -- keeping the data accessible. It would be helpful to differentiate these rules a bit more.**

**Line 21: is a good advice => is good advice**

We have fixed this typo.

Line 136: States that an example of machine readable open format that would not be easy to process include Microsoft word, but in the preceding rule (Rule 4) it states that Microsoft excel is not an open format. Perhaps just keep PDF as the example.

Lines 140-144: discussing source code doesn't really fit in this section. Perhaps move these lines to the section about metadata (Rule 7).

**Line 147: variable is a column => variable as a column**

We have fixed this typo.

**Line 147-148: unit is a table => unit as a table**

We have fixed this typo.

**Line 161: standards-compliant data, it _is_ easier to build**

We have fixed this typo.

**Line 200: meta-data in in well => meta-data in well**

We have fixed this typo

**Line 201: Extraneous close paren ) after JSON**

We have deleted this dangling parentheses.

**Line 244: Seems like an appeal to authority. Replace explicit reference to Kristen Briney**

**Line 245: Include a caution with respect to Rule 8 (ensure that the offsite "cloud" storage is as secure as the primary storage)**

**Line 311: include oxford comma "animals, fungi, and microbes"**

An oxford comma has been added.

**Line 337-338: ways to programatically query databases through**

This sentence has been restructured

**_Reviewer #1:_** _Dr. Hart and co-authors presented a comprehensive review on the essential aspects of digital data storage. While situation in reality can become very complicated, these ten simple rules can still provide good and useful guidance for researchers to develop data storage and management systems. Actually, we have pretty much followed these rules in our own practices. The manuscript was concise and well written.

Minor suggestions:_

**Page 2, Line 50: "can help" -> "can help to"**

We have have added the word to

**Page 3, Line 89: "can be difficult to impossible for others" -> "can be difficult for others"**

We have deleted the word "impossible"

**_Reviewer #3:_** This manuscript discusses ten simple rules for digital data storage. The ten rules provide simple guidelines for how researchers across a variety of disciplines can store their data for long-term access and ensure that the data are usable by others in the future. Compared with other "ten Simple Rules" papers on scientific data management and curation, this manuscript further addresses another facet of data, longer term storage best practices. The manuscript represents the collaboration among co-authors from a wide variety of academic backgrounds, who are all part of the Software Carpentry project.

**The manuscript could benefit from addressing specific needs for the storage of biological and biomedical data to be more in scope with the main themes of the journal.**

### 2016-06-01 resubmission_materials/reviewer_response.md
addressed reviewer 1 comments on line 140-144

@@
 
+ We believe data can be tidy (rule 5) but in a closed format (rule 4) or vice versa.
+While we agree with the reviewers comment that it is a similar theme, it is easy to imagine a scenario where a scientist might put very tidy data in a closed format such as Excel.
+We have renamed Rule 5 to "Data should be structured for analysis" and also substantially restructured the paragraphs of rule 5 to enhance this distinction.
+
 **Line 21: is a good advice => is good advice**
@@
 
-Line 136: States that an example of machine readable open format that would not be easy to process include Microsoft word, but in the preceding rule (Rule 4) it states that Microsoft excel is not an open format.
-Perhaps just keep PDF as the example.
+**Line 136: States that an example of machine readable open format that would not be easy to process include Microsoft word, but in the preceding rule (Rule 4) it states that Microsoft excel is not an open format.
+Perhaps just keep PDF as the example.**
 
-Lines 140-144: discussing source code doesn't really fit in this section. Perhaps move these lines to the section about metadata (Rule 7).
+We removed this line from rule 5.
+
+**Lines 140-144: discussing source code doesn't really fit in this section. Perhaps move these lines to the section about metadata (Rule 7).**
+
+We removed the reference to metadata as the reviewer notes it is probably not truly metadata but is more like provenance of an analysis.
+However we feel it is important to highlight that structured data facilitates the transparency of an analysis and have kept those lines in.
 
@@
 
-**_Reviewer #1:_** _Dr.
+**_Reviewer #2:_** Dr.
 Hart and co-authors presented a comprehensive review on the essential aspects of digital data storage.

### 2016-06-09 resubmission_materials/reviewer_response.md
Add response to Reviewer 1 point about rules 4 & 5
Finshed up last to responses to reviewer 1

@@
 
+ We have differentiated the rules by focusing on data file formats in Rule 4 and data structure in Rule 5.
 We believe data can be tidy (rule 5) but in a closed format (rule 4) or vice versa.
@@
 
+We have removed this appeal and replaced it with a citation to the blog post.
+
 **Line 245: Include a caution with respect to Rule 8 (ensure that the offsite "cloud" storage is as secure as the primary storage)**
+
+We added an additional clause that advises users to be sure that offsite storage is secure
 
@@
 
-**_Reviewer #2:_** Dr.
+**_Reviewer #2:_** _Dr.
 Hart and co-authors presented a comprehensive review on the essential aspects of digital data storage.
@@
 
-**_Reviewer #3:_** This manuscript discusses ten simple rules for digital data storage.
+**_Reviewer #3:_** _This manuscript discusses ten simple rules for digital data storage.
 The ten rules provide simple guidelines for how researchers across a variety of disciplines can store their data for long-term access and ensure that the data are usable by others in the future.
 Compared with other "ten Simple Rules" papers on scientific data management and curation, this manuscript further addresses another facet of data, longer term storage best practices.
-The manuscript represents the collaboration among co-authors from a wide variety of academic backgrounds, who are all part of the Software Carpentry project.
+The manuscript represents the collaboration among co-authors from a wide variety of academic backgrounds, who are all part of the Software Carpentry project._
 

### 2016-06-09 resubmission_materials/reviewer_response.md
responded to reviewer 3, #128

@@
 **The manuscript could benefit from addressing specific needs for the storage of biological and biomedical data to be more in scope with the main themes of the journal.**
+
+While we appreciate the reviewers suggestion, we believe that keeping the paper more generic broadens its appeal and helps future proof the paper against the rapid pace of change in the field.
+We feel any specific advice may be obsolete in a short time and therefore we prefer to keep the paper as it is.

### 2016-06-10 resubmission_materials/reviewer_response.md

--- 2016-06-09 resubmission_materials/reviewer_response.md
@@
 
-We have removed this appeal and replaced it with a citation to the blog post.
+Line 252-253: We have removed this appeal and replaced it with a citation to the blog post.
 
@@
 
-We added an additional clause that advises users to be sure that offsite storage is secure
+Line 253-254: We added an additional clause that advises users to be sure that offsite storage is secure
 
@@
 **The manuscript could benefit from addressing specific needs for the storage of biological and biomedical data to be more in scope with the main themes of the journal.**
+
+We have added several examples in our rules that are specific to biomedical data to address this concern.
+
+Line 24-26: We suggest GenBank as an example place to archive sequence data
+Line 64-66: We suggest using community tools for formatting sequence metadata
+Line 177-178: We give an example of version numbers that are specific to NCBI repositories
+
+We believe these help bring the manuscript more in line with the main themes of the journal as the author suggests.

### 2016-06-28 resubmission_materials/reviewer_response.md
Revise response to reviewers to reflect revisions
Also wrapped text.

@@
 
-Thank you for reviewing our manuscript "Ten Simple Rules for Digital Data Storage" and considering our resubmission.
-We appreciate the thoughtful feedback reviewers have provided and incorporated their comments into our revised manuscript.
-Below is a detailed accounting of our changes, with reviewer submissions highlighted in bold.
+Thank you for reviewing our manuscript "Ten Simple Rules for Digital
+Data Storage" and considering our resubmission. We appreciate the
+thoughtful feedback reviewers have provided and incorporated their
+comments into our revised manuscript. Below is a detailed accounting
+of our changes, with reviewer submissions highlighted in bold.
 
-**_Reviewer #1:_** _The manuscript enumerates a set of "rules" that promote up-front thought during experimental design regarding the long term storage of data generated.
-Each of the ten rules has a specific focus on one aspect that researchers should consider.
-The authors clearly communicate the rules with relevant supporting observations.
-A few minor issues:_
+**_Reviewer #1:_** _The manuscript enumerates a set of "rules" that
+promote up-front thought during experimental design regarding the long
+term storage of data generated. Each of the ten rules has a specific
+focus on one aspect that researchers should consider. The authors
+clearly communicate the rules with relevant supporting observations. A
+few minor issues:_
 
-**Rules 4 and 5 seem like variations on the same rule -- keeping the data accessible. It would be helpful to differentiate these rules a bit more.**
+**Rules 4 and 5 seem like variations on the same rule -- keeping the
+data accessible. It would be helpful to differentiate these rules a
+bit more.**
 
- We have differentiated the rules by focusing on data file formats in Rule 4 and data structure in Rule 5.
-We believe data can be tidy (rule 5) but in a closed format (rule 4) or vice versa.
-While we agree with the reviewers comment that it is a similar theme, it is easy to imagine a scenario where a scientist might put very tidy data in a closed format such as Excel.
-We have renamed Rule 5 to "Data should be structured for analysis" and also substantially restructured the paragraphs of rule 5 to enhance this distinction.
+ We have differentiated the rules by focusing on data file formats in
+ Rule 4 and data structure in Rule 5. We believe data can be tidy
+ (rule 5) but in a closed format (rule 4) or vice versa. While we
+ agree with the reviewers comment that these rules focus on a similar
+ theme, it is easy to imagine a scenario where a scientist might put
+ very tidy data in a closed format such as JMP or Excel files. We have
+ renamed Rule 5 to "Data should be structured for analysis" and also
+ substantially restructured the paragraphs of rule 5 to enhance this
+ distinction.
 
@@
 
-**Line 136: States that an example of machine readable open format that would not be easy to process include Microsoft word, but in the preceding rule (Rule 4) it states that Microsoft excel is not an open format.
-Perhaps just keep PDF as the example.**
+**Line 136: States that an example of machine readable open format
+that would not be easy to process include Microsoft word, but in the
+preceding rule (Rule 4) it states that Microsoft excel is not an open
+format. Perhaps just keep PDF as the example.**
 
@@
 
-**Lines 140-144: discussing source code doesn't really fit in this section. Perhaps move these lines to the section about metadata (Rule 7).**
+**Lines 140-144: discussing source code doesn't really fit in this
+section. Perhaps move these lines to the section about metadata (Rule
+7).**
 
-We removed the reference to metadata as the reviewer notes it is probably not truly metadata but is more like provenance of an analysis.
-However we feel it is important to highlight that structured data facilitates the transparency of an analysis and have kept those lines in.
+We removed the reference to metadata as the reviewer notes it is
+probably not truly metadata but is more like provenance of an
+analysis. However we feel it is important to highlight that structured
+data facilitates the transparency of an analysis and have kept those
+lines in.
 
@@
 
-**Line 244: Seems like an appeal to authority. Replace explicit reference to Kristen Briney**
+**Line 244: Seems like an appeal to authority. Replace explicit
+reference to Kristen Briney**
 
-Line 252-253: We have removed this appeal and replaced it with a citation to the blog post.
+Line 252-253: We have removed this appeal and replaced it with a
+citation to the blog post.
 
-**Line 245: Include a caution with respect to Rule 8 (ensure that the offsite "cloud" storage is as secure as the primary storage)**
+**Line 245: Include a caution with respect to Rule 8 (ensure that the
+offsite "cloud" storage is as secure as the primary storage)**
 
-Line 253-254: We added an additional clause that advises users to be sure that offsite storage is secure
+Line 253-254: We added an additional clause that advises users to be
+sure that offsite storage is secure
 
@@
 
-**_Reviewer #2:_** _Dr.
-Hart and co-authors presented a comprehensive review on the essential aspects of digital data storage.
-While situation in reality can become very complicated, these ten simple rules can still provide good and useful guidance for researchers to develop data storage and management systems.
-Actually, we have pretty much followed these rules in our own practices.
-The manuscript was concise and well written.
+**_Reviewer #2:_** _Dr. Hart and co-authors presented a comprehensive
+review on the essential aspects of digital data storage. While
+situation in reality can become very complicated, these ten simple
+rules can still provide good and useful guidance for researchers to
+develop data storage and management systems. Actually, we have pretty
+much followed these rules in our own practices. The manuscript was
+concise and well written.
 
@@
 
-**Page 3, Line 89: "can be difficult to impossible for others" -> "can be difficult for others"**
+**Page 3, Line 89: "can be difficult to impossible for others" -> "can
+be difficult for others"**
 
@@
 
-**_Reviewer #3:_** _This manuscript discusses ten simple rules for digital data storage.
-The ten rules provide simple guidelines for how researchers across a variety of disciplines can store their data for long-term access and ensure that the data are usable by others in the future.
-Compared with other "ten Simple Rules" papers on scientific data management and curation, this manuscript further addresses another facet of data, longer term storage best practices.
-The manuscript represents the collaboration among co-authors from a wide variety of academic backgrounds, who are all part of the Software Carpentry project._
+**_Reviewer #3:_** _This manuscript discusses ten simple rules for
+digital data storage. The ten rules provide simple guidelines for how
+researchers across a variety of disciplines can store their data for
+long-term access and ensure that the data are usable by others in the
+future. Compared with other "ten Simple Rules" papers on scientific
+data management and curation, this manuscript further addresses
+another facet of data, longer term storage best practices. The
+manuscript represents the collaboration among co-authors from a wide
+variety of academic backgrounds, who are all part of the Software
+Carpentry project._
 
-**The manuscript could benefit from addressing specific needs for the storage of biological and biomedical data to be more in scope with the main themes of the journal.**
+**The manuscript could benefit from addressing specific needs for the
+storage of biological and biomedical data to be more in scope with the
+main themes of the journal.**
 
-We have added several examples in our rules that are specific to biomedical data to address this concern.
+We have added several examples in our rules that are specific to
+biomedical data to address this concern. In particular:
 
-Line 24-26: We suggest GenBank as an example place to archive sequence data
-Line 64-66: We suggest using community tools for formatting sequence metadata
-Line 177-178: We give an example of version numbers that are specific to NCBI repositories
+Line 24-26: We suggest GenBank as an example place to archive sequence
+data
 
-We believe these help bring the manuscript more in line with the main themes of the journal as the author suggests.
+Line 64-66: We suggest using community tools for formatting
+sequence data and ecological metadata
+
+Line 177-178: We give an example of version numbers
+that are specific to NCBI repositories
+
+We believe these help bring the manuscript more in line with the main
+themes of the journal as the author suggests.

### 2016-07-07 resubmission_materials/reviewer_response.md
fixed long URL bug

@@
 
-Thank you for reviewing our manuscript "Ten Simple Rules for Digital
-Data Storage" and considering our resubmission. We appreciate the
-thoughtful feedback reviewers have provided and incorporated their
-comments into our revised manuscript. Below is a detailed accounting
-of our changes, with reviewer submissions highlighted in bold.
+Below is a detailed accounting of our changes, with reviewer submissions highlighted in bold.
 
@@
 much followed these rules in our own practices. The manuscript was
-concise and well written.
+concise and well written._
 
-Minor suggestions:_
+_Minor suggestions:_
 

### 2016-09-01 manuscript/manuscript.tex
rebuilt manuscript, removed tex from ignore file for final submission

@@
 % Use nameref to cite supporting information files (see Supporting Information section for more info)
-\usepackage{nameref,hyperref}
+\usepackage{nameref}
 
@@
 %\doublespacing
+
+%to handle url breaks in references
+\usepackage{url}
+\usepackage{hyperref}
+\usepackage{breakurl}
 
@@
 \newcommand{\ipsum}{{\bf IPSUM}}
+% The below needed to build markdown documents containing lists with pandoc
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}
+}
 
@@
 {\Large
-\textbf{Ten Simple Rules for Data Storage}
+\textbf{Ten Simple Rules for Digital Data Storage}
 }
@@
 Pauline Barmby $^{2 }$,
-Jeffrey Hollister $^{3 }$,
-\\
+David LeBauer $^{3 }$,
+François Michonneau $^{4,5 }$,
+Sarah Mount $^{6 }$,
+Patrick Mulrooney $^{11 }$,
+Timothée Poisot $^{7 }$,
+Kara H. Woo $^{8 }$,
+Naupaka B. Zimmerman $^{9 }$,
+Jeffrey Hollister $^{10 }$\\
 \bigskip
-\bf{1} Univeristy of Vermont, Department of Biology, Burlington
+\bf{1} University of Vermont, Department of Biology, Burlington
 \\
@@
 \\
-\bf{3} US Environmental Protection Agency, Atlantic Ecology Division
+\bf{3} University of Illinois at Urbana-Champaign, National Center for Supercomputing Applications and Institute for
+Genomic Biology
+\\
+\bf{4} University of Florida, iDigBio, Florida Museum of Natural History
+\\
+\bf{5} University of Florida, Whitney Laboratory for Marine Bioscience
+\\
+\bf{6} King's College London, Department of Informatics
+\\
+\bf{7} Université de Montréal, Département de Sciences Biologiques
+\\
+\bf{8} Washington State University, Center for Environmental Research, Education, and Outreach
+\\
+\bf{9} University of Arizona, School of Plant Sciences
+\\
+\bf{10} US Environmental Protection Agency, Atlantic Ecology Division
+\\
+\bf{11} University of California at San Diego, San Diego Supercomputer Center
 \\
 \bigskip
- $\ast$ E-mail: emh@emhart.info
+ $\ast$ E-mail: edmund.m.hart@gmail.com
 
 \end{flushleft}
-
-\section*{Abstract}\label{abstract}
-\addcontentsline{toc}{section}{Abstract}
 
@@
 
-Some example text with a citation {[}1{]}
+Data is the central currency of science, but the nature of scientific
+data has changed dramatically with the rapid pace of technology. This
+change has led to the development of a wide variety of data formats,
+dataset sizes, data complexity, data use cases, and data sharing
+practices. Improvements in high throughput DNA sequencing, sustained
+institutional support for large sensor networks {[}1,2{]}, and sky
+surveys with large-format digital cameras {[}3{]} have created massive
+quantities of data. At the same time, the combination of increasingly
+diverse research teams {[}4{]} and data aggregation in portals (e.g.~for
+biodiversity data, \href{http://gbif.org/}{GBIF.org} or
+\href{https://www.idigbio.org/portal/}{iDigBio}) necessitates increased
+coordination among data collectors and institutions {[}5,6{]}. As a
+consequence, ``data'' can now mean anything from petabytes of
+information stored in professionally-maintained databases, through
+spreadsheets on a single computer, to hand-written tables in lab
+notebooks on shelves. All remain important, but data curation practices
+must continue to keep pace with the changes brought about by new forms
+of data and new data collection and storage practices.
 
-\section*{Rule 1: Rule}\label{rule-1-rule}
-\addcontentsline{toc}{section}{Rule 1: Rule}
+While much has been written about both the virtues of data sharing
+{[}7,8{]} and the best practices to do so {[}9,10{]}, data storage has
+received comparatively less attention. Proper storage is a prerequisite
+to sharing, and indeed inadequate storage contributes to the phenomenon
+of data decay or to ``data entropy'', where data, whether publicly
+shared or not, becomes less accessible through time {[}11--14{]}. Best
+practices for data storage often begin and end with, ``deposit your data
+in a community standard repository.'' This is good advice, especially
+considering your data is most likely to be reused if it is available on
+a community site. Community repositories can also provide guidance for
+best practices. As an example, if you are archiving sequencing data, a
+repository such as those run by the National Center for Biotechnology
+Information (NCBI) (e.g.~GenBank) provides not only a location for data
+archival, but also encourages a set of practices related to consistent
+data formatting and the inclusion of appropriate metadata. However, data
+storage policies are highly variable between repositories {[}15{]}. A
+data management plan utilizing best practices across all stages of the
+data life cycle will facilitate transition from local storage to
+repository {[}16{]}. Similarly, having such a plan can facilitate
+transition from repository to repository if funding runs out or
+requirements change. Good storage practices are important even (or
+especially) in cases where data may not fit with an existing repository,
+where only derived data products (versus raw data) are suitable for
+archiving, or in the case where an existing repository may have lax
+standards.
 
-\section*{Rule 2: Rule}\label{rule-2-rule}
-\addcontentsline{toc}{section}{Rule 2: Rule}
+This manuscript describes 10 simple rules for digital data storage that
+grew out of a long discussion among instructors for the Software and
+Data Carpentry initiatives {[}17,18{]}. Software and Data Carpentry
+instructors are scientists from diverse backgrounds who have encountered
+a variety of data storage challenges and are active in teaching other
+scientists best practices for scientific computing and data management.
+Thus, this paper represents a distillation of collective experience, and
+hopefully will be useful to scientists facing a variety of data storage
+challenges. We additionally provide a glossary of common vocabulary for
+readers who may not be familiar with particular terms.
 
-\section*{Rule 3: Rule}\label{rule-3-rule}
-\addcontentsline{toc}{section}{Rule 3: Rule}
+\section*{Rule 1: Anticipate how your data will be
+used}\label{rule-1-anticipate-how-your-data-will-be-used}
+\addcontentsline{toc}{section}{Rule 1: Anticipate how your data will be
+used}
 
-\section*{Rule 4: Rule}\label{rule-4-rule}
-\addcontentsline{toc}{section}{Rule 4: Rule}
+One can avoid most of the troubles encountered during the analysis,
+management, and release of data by having a clear roadmap of what to
+expect \emph{before} data acquisition starts. For instance:
 
-\section*{Rule 5: Rule}\label{rule-5-rule}
-\addcontentsline{toc}{section}{Rule 5: Rule}
+\begin{itemize}
+\tightlist
+\item
+  How will the raw data be received? Are they delivered by a machine or
+  software, or typed-in?
+\item
+  What is the format expected by the software used for analysis?
+\item
+  Is there a community standard format for this type of data?
+\item
+  How much data will be collected, and over what period of time?
+\end{itemize}
 
-\section*{Rule 6: Rule}\label{rule-6-rule}
-\addcontentsline{toc}{section}{Rule 6: Rule}
+The answers to these questions can range from simple cases (e.g.,
+sequencing data stored in the FASTA format, which can be used ``as is''
+throughout the analysis), to experimental designs involving multiple
+instruments, each with its own output format and processing conventions.
+Knowing the state in which the data needs to be at each step of the
+analysis can help to (i) identify software tools to use in converting
+between data formats, (ii) orient technological choices about how and
+where the data should be stored, and (iii) rationalize the analysis
+pipeline, making it more amenable to re-use {[}19{]}.
 
-\section*{Rule 7: Rule}\label{rule-7-rule}
-\addcontentsline{toc}{section}{Rule 7: Rule}
+Also key is the ability to estimate the storage volume needed to store
+the data, both during and after the analysis. The required strategy will
+differ for datasets of varying size. Smaller datasets (e.g.~a few
+megabytes in size) can be managed locally with a simple data management
+plan, whereas larger datasets (e.g.~gigabytes to petabytes) will in
+almost all cases require careful planning and preparation (Rule 10).
 
-\section*{Rule 8: Rule}\label{rule-8-rule}
-\addcontentsline{toc}{section}{Rule 8: Rule}
+Lastly, early consideration and planning should be given to the metadata
+of the project. A plan should be developed early as to what metadata
+will be collected, and how it will be maintained and stored (Rule 7).
+Also be sure to consider community software tools that can facilitate
+metadata curation and repository submission. Examples in the biological
+sciences include \emph{Morpho} for ecological metadata {[}20{]} and
+\emph{mothur} {[}21{]} for submitting to NCBI's Sequence Read Archive.
 
-\section*{Rule 9: Rule}\label{rule-9-rule}
-\addcontentsline{toc}{section}{Rule 9: Rule}
+\section*{Rule 2: Know your use case}\label{rule-2-know-your-use-case}
+\addcontentsline{toc}{section}{Rule 2: Know your use case}
 
-\section*{Rule 10: Rule}\label{rule-10-rule}
-\addcontentsline{toc}{section}{Rule 10: Rule}
+Well-identified use cases make data storage easier. Ideally, prior to
+beginning data collection, researchers should be able to answer the
+following questions:
 
-\section*{Figure Legends}\label{figure-legends}
-\addcontentsline{toc}{section}{Figure Legends}
+\begin{itemize}
+\tightlist
+\item
+  Should the raw data be archived (Rule 3)?
+\item
+  Should the data used for analysis be prepared once, or re-generated
+  from the raw data each time (and what difference would this choice
+  make for storage, computing requirements, and reproducibility)?
+\item
+  Can manual corrections be avoided in favor of programmatic or
+  self-documenting approaches (e.g., Jupyter notebook or R markdown)?
+\item
+  How will changes to the data be tracked, and where will these tracked
+  changes be logged?
+\item
+  Will the final data be released, and if so, in what format?
+\item
+  Are there restrictions or privacy concerns associated with the data
+  (e.g.~for survey results with personally identifiable information
+  (PII), threatened species, or confidential business information)?
+\item
+  Will institutional validation be required prior to releasing the data?
+\item
+  Does the funding agency mandate data deposition in a publicly
+  available archive, and if so, when, where, and under what license?
+\item
+  Does the target journal mandate data deposition?
+\end{itemize}
 
-Figures here: Will need to figure out numbering\ldots{}
+None of these questions have universal answers, nor are they the only
+questions to ask before starting data acquisition. But knowing the what,
+when, and how of \emph{your} use of the data will bring you close to a
+reliable roadmap on how to handle data from acquisition through
+publication and archival.
 
-\section*{Tables}\label{tables}
-\addcontentsline{toc}{section}{Tables}
+\section*{Rule 3: Keep raw data raw}\label{rule-3-keep-raw-data-raw}
+\addcontentsline{toc}{section}{Rule 3: Keep raw data raw}
 
-Tables here: Will need to figure out numbering\ldots{}
+Since analytical and data processing procedures improve or otherwise
+change over time, having access to the `raw' (unprocessed) data can
+facilitate future re-analysis and analytical reproducibility. As
+processing algorithms improve and computational power increases, new
+analyses will be enabled that were not possible at the time of the
+original work. If only derived data are stored, it can be difficult for
+other researchers to confirm analytical results, to assess the validity
+of statistical models, or to directly compare findings across studies.
 
-\nolinenumbers
+Therefore, data should always be kept in raw format whenever possible
+(within the constraints of technical limitations). In addition to being
+the most appropriate way to ensure transparency in analysis, having the
+data stored and archived in their original state gives a common point of
+reference for derivative analyses. What constitutes sufficiently ``raw''
+data is not always clear (e.g., ohms from a temperature sensor or images
+of an Illumina sequencing flowcell are generally not archived after the
+initial processing). Yet the spirit of this rule is that data should be
+as ``pure'' as possible when they are stored. If derivations occur, they
+should be documented by also archiving relevant code and intermediate
+datasets.
+
+A cryptographic hash (e.g., SHA or MD5) of the raw data should be
+generated and distributed with the data. These hashes ensure that the
+dataset has not suffered any silent corruption/manipulation while being
+stored or transferred (see
+\href{https://www.xsede.org/news/-/news/item/6390}{Internet2 Silent Data
+Corruption}). For large enough datasets the likelihood of silent data
+corruption is high. This technique has been widely used by many Linux
+distributions to distribute images and has been very effective with
+minimal effort.
+
+\section*{Rule 4: Store data in open
+formats}\label{rule-4-store-data-in-open-formats}
+\addcontentsline{toc}{section}{Rule 4: Store data in open formats}
+
+To maximize accessibility and long-term value, it is preferable to store
+data in formats whose specifications are freely available. The
+appropriate file type will depend on the data being stored (e.g.~numeric
+measurements, text, images, video), but the key idea is that accessing
+data should not require proprietary software, hardware, or purchase of a
+commercial license. Proprietary formats change, maintaining
+organizations go out of business, and changes in license fees make
+access to data in proprietary formats unaffordable and risky for
+end-users. Examples of open data formats include comma-separated values
+(CSV) for tabular data, hierarchical data format (HDF) {[}22{]} and
+NetCDF {[}23{]} for hierarchically structured scientific data, portable
+network graphics (PNG) for images, KML (or other Open Geospatial
+Consortium (OGC) format) for spatial data, and extensible markup
+language (XML) for documents. Examples of closed formats include DWG for
+AutoCAD drawings, Photoshop document (PSD) for bitmap images, Windows
+Media Audio (WMA) for audio recording files, and Microsoft Excel (XLS)
+for tabular data. Even if day-to-day processing uses closed formats
+(e.g., due to software requirements), data being stored for archival
+purposes should be stored in open formats. This is generally not
+prohibitive; most closed-source software products enable users to export
+data to an open format.
+
+Not only should data be stored in an open format, but it should also be
+stored in a format that computers can easily use for processing. This is
+especially crucial as datasets become larger. Making data easily-usable
+is best achieved by using standard data formats that have open
+specifications (e.g., CSV, XML, JSON, HDF5), or by using databases. Such
+data formats can be handled by a variety of programming languages, as
+efficient and well-tested libraries for parsing them are typically
+available. These standard data formats also ensure interoperability,
+facilitate re-use, and reduce the chances of data loss or mistakes being
+introduced during conversion between formats. Examples of
+machine-readable open formats that would \emph{not} be easy to process
+include data included in the text of a PDF file, or scanned images of
+tabular data from a paper source.
+
+\section*{Rule 5: Data should be structured for
+analysis}\label{rule-5-data-should-be-structured-for-analysis}
+\addcontentsline{toc}{section}{Rule 5: Data should be structured for
+analysis}
+
+To take full advantage of data, it can be useful for it to be structured
+in a way that makes use, interpretation, and analysis easy. One such
+structure for data stores each variable as a column, each observation as
+a row, and each type of observational unit as a table (Fig. 1). The
+technical term for this structure is `Codd's 3rd normal form', but it
+has been made more accessible as theconcept of \emph{tidy} data
+{[}24{]}. When data is organized in this way, the duplication of
+information is reduced and it is easier to subset or summarize the
+dataset to include the variables or observations of interest.
+
+\textbf{Figure 1}: Example of an untidy dataset (A) and its tidy
+equivalent (B). Dataset A is untidy because it mixes observational units
+(species, location of observations, measurements about individuals), the
+units are mixed and listed with the observations, more than one variable
+is listed (both latitude and longitude for the coordinates, and genus
+and species for the species names), and several formats are used in the
+same column for dates and geographic coordinates. Dataset B is an
+example of a tidy version of dataset A that reduces the amount of
+information that is duplicated in each row, limiting chances of
+introducing mistakes in the data. By having species in a separate table,
+they can be identified uniquely using the Taxonomic Serial Number (TSN)
+from the Integrated Taxonomic Information System (ITIS), and it makes it
+easy to add information about the classification of these species. It
+also allows researchers to edit the taxonomic information independently
+from the table that holds the measurements about the individuals. Unique
+values for each observational unit facilitate the programmatic
+combination of information using ``join'' operations. With this example,
+if the focus of the study for which these data were collected is based
+upon the size measurements of the individuals (weight and length),
+information about ``where'', ``when'', and ``what'' animals were
+measured can be considered meta-data. Using the tidy format makes this
+distinction clearer.
+
+One axiom about the structure of data and code holds that one should
+``write code for humans, write data for computers'' {[}25{]}. When data
+can be easily imported and manipulated using familiar software, whether
+via a scripting language, a spreadsheet, or any other computer program
+that can import these common files, data becomes easier to re-use.
+Furthermore, having the source code for the software doing the analysis
+available provides provenance for how the data is processed and
+analyzed. This makes analysis more transparent, since all assumptions
+about the structure of the data are implicitly stated in the source
+code. This also enables extraction of the analyses performed, their
+reproduction, and their modification.
+
+Interoperability is facilitated when variable names are mapped to
+existing data standards. For instance, for biodiversity data, the
+\href{http://www.tdwg.org/standards/450/}{Darwin Core Standard} provides
+a set of terms that describe observations, specimens, samples, and
+related information for a taxa. For earth science and ecosystem models
+and data, the \href{https://cfconventions.org}{Climate Forecasting
+Conventions} are widely adopted, such that a large ecosystem of software
+and data products exist to reduce the technical burden of reformatting
+and reusing large and complex data. Because each term in such standards
+is clearly defined and documented, each dataset can use the terms
+consistently; this facilitates data sharing across institutions,
+applications, and disciplines. With machine-readable,
+standards-compliant data, it becomes easier to build an Application
+Programming Interface (API) to query the dataset and retrieve a subset
+of interest, as outlined in Rule 10.
+
+\section*{Rule 6: Data should be uniquely
+identifiable}\label{rule-6-data-should-be-uniquely-identifiable}
+\addcontentsline{toc}{section}{Rule 6: Data should be uniquely
+identifiable}
+
+To aid reproducibility, the data used in a scientific publication should
+be uniquely identifiable. Ideally, datasets should have a unique
+identifier such as a Digital Object Identifier (DOI), Archival Resource
+Key (ARK), or a persistent URL (PURL). An increasing number of online
+services, such as \href{http://figshare.com/}{Figshare},
+\href{http://zenodo.org}{Zenodo}, or
+\href{http://www.dataone.org}{DataOne}, are able to provide these.
+Institutional initiatives also exist, and are known to your local
+librarians. Some repositories may require specific identifiers and these
+could change with time. For instance, NCBI sequence data will in the
+future only be identified by ``accession.version'' IDs. The ``GI''
+identifiers (in use since 1994) will be retired in late 2016 {[}26{]}.
+
+Even as identifier standards may change over time, datasets can evolve
+over time as well. In order to distinguish between different versions of
+the same data, each dataset should have a distinct name, which includes
+a version identifier. A simple way to do this is to use date stamps as
+part of the dataset name. Using the ISO 8601 standard avoids regional
+ambiguities: it mandates the date format \texttt{YYYY-MM-DD} (i.e.~from
+largest time unit to smallest). For example, the date ``February 1st,
+2015'', while written as \texttt{01-02-2015} in the UK and
+\texttt{02-01-2015} in the US, is unambiguous (\texttt{2015-02-01})
+under this standard.
+
+Semantic versioning is a richer approach to solving the same problem
+{[}27{]}. The CellPack datasets are an example of this {[}28{]}. A
+semantic version number takes the form: \texttt{Major.Minor.Patch}, e.g.
+\texttt{0.2.7}. The \emph{major version} numbers should be incremented
+(or bumped) when a dataset scheme has been updated, or some other change
+is made that is not compatible with previous versions of the data with
+the same major version number. This means that an experiment using
+version \texttt{1.0.0} of the dataset may not run on version
+\texttt{2.0.0} without changes to the data analysis. The \emph{minor
+version} should be bumped when a change has been made which is
+compatible with older versions of the data with the same major version.
+This means that any analysis that can be performed on version
+\texttt{1.0.0} of the data is repeatable with version \texttt{1.1.0} of
+the data. For example, adding a new year in a temporal survey will
+result in a bump in the minor version. The \emph{patch version} number
+is bumped when typos or bugs have been fixed. For example version
+\texttt{1.0.1} of a dataset may fix a typo in version \texttt{1.0.0}.
+
+\section*{Rule 7: Link relevant
+metadata}\label{rule-7-link-relevant-metadata}
+\addcontentsline{toc}{section}{Rule 7: Link relevant metadata}
+
+Metadata is the contextual information required to interpret data (Fig.
+1) and should be clearly defined and tightly integrated with data. The
+importance of metadata for context, reusability, and discovery has been
+written about at length in guides for data management best practices
+{[}9,13,29{]}.
+
+Metadata should be as comprehensive as possible, using standards and
+conventions of a discipline, and should be machine-readable. Metadata
+should always accompany a dataset, wherever it is stored, but the best
+way to do this depends on the format of the data. Text files can contain
+meta-data in well-defined text files such as XML or JSON. Some file
+formats are self-documenting; for example NetCDF, HDF5, and many image
+file formats allow for embedded metadata {[}22,23{]}. In a relational
+database, metadata tables can be clearly labeled and linked to the data.
+Ideally, a schema will be provided that also shows the linkages between
+data tables and metadata tables. Another--simpler--scenario is a set of
+flat (non-hierarchical) text files--in this case a semantically
+versioned, compressed archive should be created that includes metadata.
+
+Whatever format is used for archiving, the goal should be to make the
+link between metadata and data as clear as possible. The best approach
+is dependent on the archiving plan used, but even if the dataset is
+archived solely for personal use, metadata will provide crucial context
+for future reuse.
+
+\section*{Rule 8: Adopt the proper privacy
+protocols}\label{rule-8-adopt-the-proper-privacy-protocols}
+\addcontentsline{toc}{section}{Rule 8: Adopt the proper privacy
+protocols}
+
+In datasets where privacy is important, be sure to have a plan in place
+to protect data confidentiality. You should consider the different data
+stakeholders when developing privacy protocols for your data storage.
+These stakeholders include funding agencies, human subjects or entities,
+collaborators, and yourself. Both the NSF and NIH have data sharing
+policies in their grant guidelines to prevent sharing personally
+identifiable information, and to anonymize data on human subjects.
+
+In small datasets, a lookup table (protecting PII by removing it and
+replacing it with a unique id that maps to the sensitive data in an
+external dataset) is enough to anonymize a minimal amount of personal
+information. Hashing techniques are susceptible to a number of attacks,
+and all hashed data can eventually be determined. Famously, New York
+City officials shared what they thought was anonymized data on cab
+drivers and over 173 million cab rides. However, it was quickly
+recognized that the city anonymized the data with a simple MD5 hashing
+scheme and all 20 GB of data were de-anonymized in a matter of hours
+{[}30{]}. This type of error can be prevented by asking a trusted
+colleague or security personnel to try to ``crack'' anonymised data
+before releasing it publicly. Often the person who has produced the data
+is not well placed to check the fine details of their own security
+procedures. If possible, the best solution is to remove any sensitive
+data that is not required from the dataset prior to distribution.
+
+In more problematic cases, the data itself allows identifiability: this
+is the case with human genomic data that map directly onto a subject's
+identity {[}31{]}. Methods for dealing with these complex issues at the
+intersection of data storage and privacy are rapidly evolving, and
+include storing changes against a reference genome to help with privacy
+and reduce overall data volumes {[}32,33{]}, and/or bringing computation
+to data storage facilities instead of vice versa {[}34{]}. Having a plan
+for privacy before data is acquired is important, because it can
+determine or limit how data will be stored.
+
+\section*{Rule 9: Have a systematic backup
+scheme}\label{rule-9-have-a-systematic-backup-scheme}
+\addcontentsline{toc}{section}{Rule 9: Have a systematic backup scheme}
+
+Every storage medium can fail, and every failure can result in loss of
+data. Researchers should therefore back data up at all stages of the
+research process. Data stored on local computers or institutional
+servers during the collection and analysis phases of a project should be
+backed up to other locations to protect against data loss. No backup
+system is failsafe (see the stories of the
+\href{https://www.insidehighered.com/news/2014/05/16/dedoose-crash-shows-dangers-handing-data-cloud-services}{Dedoose
+crash} and the
+\href{http://thenextweb.com/media/2012/05/21/how-pixars-toy-story-2-was-deleted-twice-once-by-technology-and-again-for-its-own-good/}{near
+deletion of Toy Story 2}), so more than one backup system should be
+used. Ideally you should have two on-site copies (such as on a computer,
+an external hard drive, or a tape) and one off-site copy (e.g.~cloud
+storage){[}35{]}, with care taken to ensure that the off-site copy is as
+secure as the on-site copies. Keeping backups in multiple locations
+additionally protects against data loss due to theft or natural
+disasters.
+
+Researchers should test their backups regularly to ensure that they are
+functioning properly. Common reasons for backup failure include:
+
+\begin{itemize}
+\tightlist
+\item
+  faulty backup software
+\item
+  incorrect configuration (e.g., not backing up sub-directories)
+\item
+  encryption (e.g., someone encrypted the backups but later lost the
+  password to decrypt them)
+\item
+  media errors
+\end{itemize}
+
+Consider the backup plan of your selected data repository before
+publishing your data and if possible, find out about the long-term
+storage plans of the repository. Many repositories mirror the data they
+host on multiple machines. Are there plans in place to keep data
+available if the organization that manages the repository dissolves?
+
+\section*{Rule 10: The location and method of data storage depend on how
+much data you
+have}\label{rule-10-the-location-and-method-of-data-storage-depend-on-how-much-data-you-have}
+\addcontentsline{toc}{section}{Rule 10: The location and method of data
+storage depend on how much data you have}
+
+The storage method you should choose depends on the size and nature of
+your data, the cost of storage and access over time, the time it takes
+to transfer the data, how the data will be used, and any privacy
+concerns. Data is increasingly generated in the range of many terabytes
+by environmental sensors, satellites, automated analytical tools,
+simulation models, and nucleic acid sequencers. Even larger data
+generating machines, like the Large Hadron Collider (LHC) and the Large
+Scale Synoptic Survey Telescope (LSST), generate many terabytes (TB) per
+day, rapidly accumulating to petabyte (PB) scale over the course of any
+particular study. While the cost of storage continues to decrease, the
+volume of data to be stored impacts the choice of storage methods and
+locations: for large datasets it is necessary to balance the cost of
+storage with the time of access and costs of re-generating the data.
+With new commercial cloud offerings (e.g., Amazon S3) the cost of
+retrieving the data might exceed the cost of analysis or of
+re-generating the data from scratch.
+
+When data takes too long to transfer or is costly to store, it can
+become more efficient to use a computer system for analysis that can
+directly access and use the data in place instead of first transferring
+it to a local machine. Inactive data can be put in longer-term storage;
+this is less expensive, but can take longer to retrieve. Some storage
+systems automatically migrate `stale' files to longer term storage.
+Alternatively, some computing can be done `in the database' or `on disk'
+via database query languages (e.g.~SQL, MapReduce) that perform basic
+arithmetic, or via the use of procedural languages (e.g.~R, Python, C)
+embedded in the database server. Modern database technologies such as
+HDFS and Spark allow these computations to be done on data of almost any
+size. When data is larger than locally available RAM, it can be handled
+by conducting analyses on a `big memory' node, which most
+high-performance computing centers have deployed. Relying on tight
+software/hardware integration, these can allow for the analysis of
+datasets around 1-4 TB in size. This allows the user to read in and use
+a large dataset without special tools.
+
+If you regularly only need access to a small subset of your data or need
+to share it with many collaborators, a web-based API (Application
+Programming Interface) might be a good solution. Using this method, many
+users can send requests to a web service which can subset the data,
+perform in-database computation, and return smaller volumes of data as
+specific slices. Tools based on web services make it easier to find and
+download data, and facilitate analysis via reproducible scripts.
+However, they can also lead to excessive and careless abuse of resources
+without proper safeguards in place. The time required to re-download and
+recompute results can be reduced by `caching'. Caching stores copies of
+downloads and generated files that are recognized when the same script
+is run multiple times.
+
+\section*{Further Reading and
+Resources}\label{further-reading-and-resources}
+\addcontentsline{toc}{section}{Further Reading and Resources}
+
+Digital data storage is a vast topic; the references given here and
+elsewhere in this paper provide some starting points for interested
+readers. For beginning users of scientific data,
+\href{http://datacarpentry.org}{Data Carpentry} offers workshops and
+resources on data management and analysis, as do the DataONE education
+modules {[}36{]}. For librarians and others who are responsible for data
+archiving, Data Curation Profiles {[}37{]} may be of interest.
+
+\section*{Glossary used in the
+manuscript}\label{glossary-used-in-the-manuscript}
+\addcontentsline{toc}{section}{Glossary used in the manuscript}
+
+\subsection*{Projects and Initiatives}\label{projects-and-initiatives}
+\addcontentsline{toc}{subsection}{Projects and Initiatives}
+
+\begin{itemize}
+\item
+  \textbf{Global Biodiversity Information Facility} (GBIF,
+  \url{http://www.gbif.org}) provides an international open data
+  infrastructure to publish and disseminate biodiversity information.
+\item
+  \textbf{Integrated Digitized Biocollections} (iDigBio,
+  \url{https://www.idigbio.org}) is a project funded by the National
+  Science Foundation that facilitates the digitization of natural
+  history collections, and provides data and images for biological
+  specimens.
+\item
+  \textbf{Integrated Taxonomic Information System} (ITIS,
+  \url{http://www.itis.gov}) is an international partnership of
+  governmental organizations that aims at providing authoritative
+  taxonomic information for plants, animals, fungi, and microbes.
+\end{itemize}
+
+\subsection*{File formats}\label{file-formats}
+\addcontentsline{toc}{subsection}{File formats}
+
+\begin{itemize}
+\item
+  \textbf{Comma-Separated Values} (CSV) and \textbf{Tab-Separated
+  Values} (TSV) are plain text file formats used to store tabular data
+  where each row is represented by a line in the file, and each field
+  (column) is separated by a comma (for CSV) or by the Tab character
+  (for TSV).
+\item
+  \textbf{FASTA} is a simple and widely used file format used to
+  represent sequences of nucleotides or amino-acids in plain text making
+  it easy to manipulate programmatically.
+\item
+  \textbf{Hierarchical Data Format} (HDF5) is an open-source binary file
+  format designed to store large amounts of data (and their associated
+  metadata) by providing a hierarchical structure that could be compared
+  to how a hard drive is organized with directories and files. It is
+  maintained by the non-profit HDF Group, a spin off of the National
+  Center for Supercomputing Applications (NCSA).
+\item
+  \textbf{JavaScript Object Notation} (JSON) is a plain text file format
+  typically used to store arbitrarily structured data in the form of
+  keys and values. It can be used to store non-relational databases as
+  it does not rely on a tabular data format. In many respects, it has
+  been replacing XML.
+\item
+  \textbf{Network Common Data Form} (NetCDF) is an open-source binary
+  file format designed to store large datasets in array-oriented
+  scientific data as typically used in the geosciences. It is maintained
+  by Unidata, a non-profit member of the University Corporation for
+  Atmospheric Research (UCAR) which is funded by the National Science
+  Foundation.
+\item
+  \textbf{Extensible Markup Language} (XML) is a markup language and the
+  file format used to store documents written with it. It is used to
+  represent arbitrary data structures and is both human and
+  machine-readable.
+\end{itemize}
+
+\subsection*{Programming and
+algorithms}\label{programming-and-algorithms}
+\addcontentsline{toc}{subsection}{Programming and algorithms}
+
+\begin{itemize}
+\item
+  \textbf{Web Application Programming Interface} (API) provide ways to
+  programmatically query databases through the internet. They notably
+  allow users to retrieve and work with a small slice of a large
+  dataset.
+\item
+  \textbf{Hadoop Distributed File System} (HDFS) is a Java-based file
+  system where data is stored in small chunks across multiple redundant
+  nodes.
+\item
+  \textbf{MapReduce} is a style of programming designed to work with
+  large datasets in parallel computing environments. Such programs are
+  composed of a \textbf{map} procedure where the dataset is sliced into
+  several pieces, and a \textbf{reduce} procedure where summary
+  operations are then applied to each of the slices.
+\item
+  \textbf{Secure Hash Algorithm 2} (SHA-2) is a family of Secure Hashing
+  Algorithms used in cryptographic analysis, often to verify the
+  integrity of a file. A cryptographic hash function converts a
+  ``message'' (e.g., passwords, file content) into an encrypted value.
+  Cryptographic hash functions are easy to compute from the message, but
+  it should be impossible to recover the message from the output, and
+  any modifications to the message should also modify the output. The
+  SHA algorithms are often used in preference to similar tools such as
+  MD5 (mentioned in Rule 8), which are no longer secure. All hashing
+  algorithms are vulnerable to brute force attacks. Key Derivation
+  Function (KDF) implementations like BCrypt and PBKDF2 are considered
+  significantly more secure, but by design more costly to compute.
+\item
+  Apache \textbf{Spark} is an open source computing platform for
+  querying large data sets in memory, in contrast to on-disk based
+  methods like MapReduce.
+\item
+  \textbf{Structured Query Language} (SQL) is a programming language
+  used to interact with relational database management systems.
+\end{itemize}
+
+\subsection*{Hardware}\label{hardware}
+\addcontentsline{toc}{subsection}{Hardware}
+
+\begin{itemize}
+\tightlist
+\item
+  \textbf{mega-, giga-, tera-, peta-bytes} are units of digital
+  information and are used to measure the size of datasets or the
+  storage media. Originally a byte was the minimum amount of memory
+  needed to store a single character of text in a computer. The prefixes
+  mega-, giga-, tera- and peta- refer to the international system of
+  units for the multiple of the unit, and correspond to \(10^{6}\),
+  \(10^{9}\), \(10^{12}\), and \(10^{15}\) abbreviated M, G, T and P
+  respectively.
+\end{itemize}
+
+\subsection*{Persistent identifiers}\label{persistent-identifiers}
+\addcontentsline{toc}{subsection}{Persistent identifiers}
+
+\begin{itemize}
+\item
+  \textbf{Archival Resource Key} (ARK) identifiers are URLs designed to
+  support long-term access to information online.
+\item
+  \textbf{Digital Object Identifier} (DOI) provides unique and
+  persistent identifiers for electronic documents (in particular journal
+  articles and datasets) on the internet. The uniqueness of the
+  identifiers is guaranteed by a central registry. By dissociating the
+  identifier and the location of the document (i.e., the URL), the DOI
+  can remain fixed even if the location of the digital object it is
+  pointing to changes.
+\item
+  \textbf{Persistent Uniform Resource Locator} (PURL) is an URL used to
+  redirect to the location of an electronic object on the internet. DOI
+  and ARK are examples of implementations of PURL.
+\item
+  \textbf{Uniform Resource Locator} (URL) gives the location of an
+  object on the World Wide Web; the most familiar type of URL is a
+  website address.
+\end{itemize}
+
+\section*{Acknowledgements}\label{acknowledgements}
+\addcontentsline{toc}{section}{Acknowledgements}
+
+We would like to thank G. Wilson and the Software Carpentry instructor
+community for providing the initial inspiration for this paper and the
+discussion held at \url{https://github.com/swcarpentry/site/issues/797}.
+We thank K. Hinsen, G. Pugh, and W. Turner for comments on the draft
+manuscript. Contributions from JWH, an employee of the U.S.
+Environmental Protection Agency (EPA) have not been subjected to the
+Agency's peer and administrative review. Therefore, the conclusions and
+opinions drawn are solely those of the author(s) and are not necessarily
+the views of the Agency. Mention of trade names or commercial products
+does not constitute endorsement or recommendation for use. SM would like
+to thank Dr.~Laurence Tratt for reviewing an early draft of this paper.
+
+\newpage
 
@@
 
-1. Goodman A, Pepe A, Blocker AW, Borgman CL, Cranmer K, Crosas M, et
+\hyperdef{}{ref-Reid2014}{\label{ref-Reid2014}}
+1. Reid JG, Carroll A, Veeraraghavan N, Dahdouli M, Sundquist A, English
+A, et al. Launching genomics into the cloud: deployment of Mercury, a
+next generation sequence analysis pipeline. BMC bioinformatics. 2014;15:
+30.
+\href{http://doi.org/10.1186/1471-2105-15-30}{doi:10.1186/1471-2105-15-30}
+
+\hyperdef{}{ref-Hampton2013}{\label{ref-Hampton2013}}
+2. Hampton SE, Strasser C, Tewksbury JJ, Gram WK, Budden AE, Batcheller
+AL, et al. Big data and the future of ecology. Frontiers in Ecology and
+the Environment. 2013; 130312142848005.
+\href{http://doi.org/10.1890/120103}{doi:10.1890/120103}
+
+\hyperdef{}{ref-Eisenstein2011}{\label{ref-Eisenstein2011}}
+3. Eisenstein DJ, others. SDSS-III: Massive Spectroscopic Surveys of the
+Distant Universe, the Milky Way, and Extra-Solar Planetary Systems. The
+Astronomical Journal. 2011;142: 72.
+
+\hyperdef{}{ref-Adams2012}{\label{ref-Adams2012}}
+4. Adams J. Collaborations: The rise of research networks. Nature.
+Nature Publishing Group, a division of Macmillan Publishers Limited. All
+Rights Reserved. 2012;490: 335--6.
+\href{http://doi.org/10.1038/490335a}{doi:10.1038/490335a}
+
+\hyperdef{}{ref-Fraser2013}{\label{ref-Fraser2013}}
+5. Fraser LH, Henry HA, Carlyle CN, White SR, Beierkuhnlein C, Cahill
+JF, et al. Coordinated distributed experiments: an emerging tool for
+testing global hypotheses in ecology and environmental science.
+Frontiers in Ecology and the Environment. Ecological Society of America;
+2013;11: 147--155.
+\href{http://doi.org/10.1890/110279}{doi:10.1890/110279}
+
+\hyperdef{}{ref-Robertson2014}{\label{ref-Robertson2014}}
+6. Robertson MAG Tim AND Döring. The GBIF integrated publishing toolkit:
+Facilitating the efficient publishing of biodiversity data on the
+internet. PLoS ONE. Public Library of Science; 2014;9: e102623.
+\href{http://doi.org/10.1371/journal.pone.0102623}{doi:10.1371/journal.pone.0102623}
+
+\hyperdef{}{ref-Wolkovich2012}{\label{ref-Wolkovich2012}}
+7. Wolkovich EM, Regetz J, O'Connor MI. Advances in global change
+research require open science by individual researchers. Global Change
+Biology. 2012;18: 2102--2110.
+\href{http://doi.org/10.1111/j.1365-2486.2012.02693.x}{doi:10.1111/j.1365-2486.2012.02693.x}
+
+\hyperdef{}{ref-Roche2014}{\label{ref-Roche2014}}
+8. Roche DG, Lanfear R, Binning SA, Haff TM, Schwanz LE, Cain KE, et al.
+Troubleshooting public data archiving: suggestions to increase
+participation. PLoS biology. 2014;12: e1001779.
+\href{http://doi.org/10.1371/journal.pbio.1001779}{doi:10.1371/journal.pbio.1001779}
+
+\hyperdef{}{ref-White2013}{\label{ref-White2013}}
+9. White E, Baldridge E, Brym Z, Locey K, McGlinn D, Supp S. Nine simple
+ways to make it easier to (re)use your data. Ideas in Ecology and
+Evolution. 2013;6: 1--10.
+\href{http://doi.org/10.4033/iee.2013.6b.6.f}{doi:10.4033/iee.2013.6b.6.f}
+
+\hyperdef{}{ref-Goodman2014}{\label{ref-Goodman2014}}
+10. Goodman A, Pepe A, Blocker AW, Borgman CL, Cranmer K, Crosas M, et
 al. Ten simple rules for the care and feeding of scientific data. PLoS
-computational biology. Public Library of Science; 2014;10: e1003542.
+computational biology. 2014;10: e1003542.
+\href{http://doi.org/10.1371/journal.pcbi.1003542}{doi:10.1371/journal.pcbi.1003542}
+
+\hyperdef{}{ref-Pepe2014}{\label{ref-Pepe2014}}
+11. Pepe A, Goodman A, Muench A, Crosas M, Erdmann C. How Do Astronomers
+Share Data? Reliability and Persistence of Datasets Linked in AAS
+Publications and a Qualitative Study of Data Practices among US
+Astronomers. Golden AA-J, editor. PLoS ONE. 2014;9: e104798.
+\href{http://doi.org/10.1371/journal.pone.0104798}{doi:10.1371/journal.pone.0104798}
+
+\hyperdef{}{ref-Vines2014}{\label{ref-Vines2014}}
+12. Vines TH, Albert AYK, Andrew RL, Débarre F, Bock DG, Franklin MT, et
+al. The availability of research data declines rapidly with article age.
+Current biology : CB. Elsevier; 2014;24: 94--7.
+\href{http://doi.org/10.1016/j.cub.2013.11.014}{doi:10.1016/j.cub.2013.11.014}
+
+\hyperdef{}{ref-Michener2012}{\label{ref-Michener2012}}
+13. Michener WK, Jones MB. Ecoinformatics: supporting ecology as a
+data-intensive science. Trends in ecology \& evolution. 2012;27: 85--93.
+\href{http://doi.org/10.1016/j.tree.2011.11.016}{doi:10.1016/j.tree.2011.11.016}
+
+\hyperdef{}{ref-Michener1997}{\label{ref-Michener1997}}
+14. Michener WK, Brunt JW, Helly JJ, Kirchner TB, Stafford SG.
+Nongeospatial metadata for the ecological sciences. Ecological
+Applications. Eco Soc America; 1997;7: 330--342.
+
+\hyperdef{}{ref-Marcial2010}{\label{ref-Marcial2010}}
+15. Marcial LH, Hemminger BM. Scientific data repositories on the Web:
+An initial survey. Journal of the American Society for Information
+Science and Technology. 2010;61: 2029--2048.
+\href{http://doi.org/10.1002/asi.21339}{doi:10.1002/asi.21339}
+
+\hyperdef{}{ref-Michener2015}{\label{ref-Michener2015}}
+16. Michener WK. Ten simple rules for creating a good data management
+plan. PLoS computational biology. Public Library of Science; 2015;11:
+e1004525.
+
+\hyperdef{}{ref-Wilson2014}{\label{ref-Wilson2014}}
+17. Wilson G. Software Carpentry: lessons learned. F1000Research.
+2014;3: 62.
+\href{http://doi.org/10.12688/f1000research.3-62.v1}{doi:10.12688/f1000research.3-62.v1}
+
+\hyperdef{}{ref-Teal2015}{\label{ref-Teal2015}}
+18. Teal TK, Cranston Ka, Lapp H, White E, Wilson G, Ram K, et al. Data
+Carpentry: Workshops to Increase Data Literacy for Researchers.
+International Journal of Digital Curation. 2015;10: 135--143.
+\href{http://doi.org/10.2218/ijdc.v10i1.351}{doi:10.2218/ijdc.v10i1.351}
+
+\hyperdef{}{ref-Roche2015}{\label{ref-Roche2015}}
+19. Roche DG, Kruuk LEB, Lanfear R, Binning SA. Public Data Archiving in
+Ecology and Evolution: How Well Are We Doing? PLoS Biology. 2015;13:
+e1002295.
+\href{http://doi.org/10.1371/journal.pbio.1002295}{doi:10.1371/journal.pbio.1002295}
+
+\hyperdef{}{ref-higgins2002}{\label{ref-higgins2002}}
+20. Higgins D, Berkley C, Jones M. Managing heterogeneous ecological
+data using morpho {[}Internet{]}. IEEE Comput. Soc; 2002. pp. 69--76.
+\href{http://doi.org/10.1109/SSDM.2002.1029707}{doi:10.1109/SSDM.2002.1029707}
+
+\hyperdef{}{ref-schloss2009}{\label{ref-schloss2009}}
+21. Schloss PD, Westcott SL, Ryabin T, Hall JR, Hartmann M, Hollister
+EB, et al. Introducing mothur: Open-source, platform-independent,
+community-supported software for describing and comparing microbial
+communities. Applied and environmental microbiology. Am Soc Microbiol;
+2009;75: 7537--7541.
+
+\hyperdef{}{ref-koziol1998hdf5}{\label{ref-koziol1998hdf5}}
+22. Koziol Q, Matzke R. Hdf5: A new generation of hdf: Reference manual
+and user guide. National Center for Supercomputing Applications,
+Champaign, Illinois, USA, http://hdf ncsa uiuc edu/nra/HDF5. 1998;
+
+\hyperdef{}{ref-rew1990netcdf}{\label{ref-rew1990netcdf}}
+23. Rew R, Davis G. NetCDF: An interface for scientific data access.
+Computer Graphics and Applications, IEEE. IEEE; 1990;10: 76--82.
+
+\hyperdef{}{ref-Wickham2014tidy}{\label{ref-Wickham2014tidy}}
+24. Wickham H. Tidy Data. Journal of Statistical Software. 2014;59:
+1--23. Available: \url{http://www.jstatsoft.org/v59/i10}
+
+\hyperdef{}{ref-buffalo2015bioinformatics}{\label{ref-buffalo2015bioinformatics}}
+25. Buffalo V. Bioinformatics data skills: Reproducible and robust
+research with open source tools. O'Reilly Media, Inc. 2015.
+
+\hyperdef{}{ref-ncbi}{\label{ref-ncbi}}
+26. NCBI. NCBI is phasing out sequence gIs - use accession.Version
+instead!
+\url{http://www.ncbi.nlm.nih.gov/news/03-02-2016-phase-out-of-GI-numbers/};
+
+\hyperdef{}{ref-semver2014}{\label{ref-semver2014}}
+27. Preston-Werner T. Semantic Versioning 2.0.0.
+\url{http://semver.org}; 2014.
+
+\hyperdef{}{ref-johnson2014cellpack}{\label{ref-johnson2014cellpack}}
+28. Johnson GT, Goodsell DS, Autin L, Forli S, Sanner MF, Olson AJ. 3D
+molecular models of whole HIV-1 virions generated with cellPACK. Faraday
+Discuss. The Royal Society of Chemistry; 2014;169: 23--44.
+\href{http://doi.org/10.1039/C4FD00017J}{doi:10.1039/C4FD00017J}
+
+\hyperdef{}{ref-Strasser2012}{\label{ref-Strasser2012}}
+29. Strasser C, Cook R, Michener W, Budden A. Primer on Data Management
+: What you always wanted to know {[}Internet{]}. California Digital
+Libraries; 2012.
+\href{http://doi.org/10.5060/D2251G48}{doi:10.5060/D2251G48}
+
+\hyperdef{}{ref-Goodin}{\label{ref-Goodin}}
+30. Goodin D. Poorly anonymized logs reveal NYC cab drivers' detailed
+whereabouts {[}Internet{]}. 2015. Available:
+\url{http://arstechnica.com/tech-policy/2014/06/poorly-anonymized-logs-reveal-nyc-cab-drivers-detailed-whereabouts/}
+
+\hyperdef{}{ref-Homer2008}{\label{ref-Homer2008}}
+31. Homer N, Szelinger S, Redman M, Duggan D, Tembe W, Muehling J, et
+al. Resolving individuals contributing trace amounts of DNA to highly
+complex mixtures using high-density SNP genotyping microarrays. PLoS
+genetics. Public Library of Science; 2008;4: e1000167.
+\href{http://doi.org/10.1371/journal.pgen.1000167}{doi:10.1371/journal.pgen.1000167}
+
+\hyperdef{}{ref-Kahn2011}{\label{ref-Kahn2011}}
+32. Kahn SD. On the future of genomic data. Science (New York, NY).
+2011;331: 728--9.
+\href{http://doi.org/10.1126/science.1197891}{doi:10.1126/science.1197891}
+
+\hyperdef{}{ref-wandelt2014trends}{\label{ref-wandelt2014trends}}
+33. Wandelt S, Bux M, Leser U. Trends in genome compression. Current
+Bioinformatics. Bentham Science Publishers; 2014;9: 315--326.
+
+\hyperdef{}{ref-Gaye2014}{\label{ref-Gaye2014}}
+34. Gaye A, Marcon Y, Isaeva J, LaFlamme P, Turner A, Jones EM, et al.
+DataSHIELD: taking the analysis to the data, not the data to the
+analysis. International journal of epidemiology. 2014;43: 1929--44.
+\href{http://doi.org/10.1093/ije/dyu188}{doi:10.1093/ije/dyu188}
+
+\hyperdef{}{ref-briney2013}{\label{ref-briney2013}}
+35. Briney K. Rule of 3. DATA AB INITIO.
+\url{http://dataabinitio.com/?p=320}; 2013.
+
+\hyperdef{}{ref-Dataone2012}{\label{ref-Dataone2012}}
+36. Henkel H, Hutchison V, Strasser C, Rebich Hespanha S, Vanderbilt K,
+Wayne L, et al. DataONE Education Modules.
+\url{https://www.dataone.org/education-modules}; 2012.
+
+\hyperdef{}{ref-Witt2009}{\label{ref-Witt2009}}
+37. Witt M, Carlson J, Brandt S, Cragin M. Constructing Data Curation
+Profiles. International Journal of Digital Curation. 2009;4: 93--103.
 

### 2016-10-01 manuscript/manuscript.tex
Built corrected proof PDF

@@
 \\
-Edmund Hart $^{1 ,\ast}$,
+Edmund M. Hart $^{1 ,\ast}$,
 Pauline Barmby $^{2 }$,
@@
 Naupaka B. Zimmerman $^{9 }$,
-Jeffrey Hollister $^{10 }$\\
+Jeffrey W. Hollister $^{10 }$\\
 \bigskip
@@
 dataset sizes, data complexity, data use cases, and data sharing
-practices. Improvements in high throughput DNA sequencing, sustained
+practices. Improvements in high-throughput DNA sequencing, sustained
 institutional support for large sensor networks {[}1,2{]}, and sky
@@
 quantities of data. At the same time, the combination of increasingly
-diverse research teams {[}4{]} and data aggregation in portals (e.g.~for
-biodiversity data, \href{http://gbif.org/}{GBIF.org} or
+diverse research teams {[}4{]} and data aggregation in portals (e.g.,
+for biodiversity data, \href{http://gbif.org/}{GBIF.org} or
 \href{https://www.idigbio.org/portal/}{iDigBio}) necessitates increased
@@
 consequence, ``data'' can now mean anything from petabytes of
-information stored in professionally-maintained databases, through
-spreadsheets on a single computer, to hand-written tables in lab
+information stored in professionally maintained databases, to
+spreadsheets on a single computer, to handwritten tables in lab
 notebooks on shelves. All remain important, but data curation practices
@@
 to sharing, and indeed inadequate storage contributes to the phenomenon
-of data decay or to ``data entropy'', where data, whether publicly
+of data decay or to ``data entropy,'' in which data, whether publicly
 shared or not, becomes less accessible through time {[}11--14{]}. Best
-practices for data storage often begin and end with, ``deposit your data
-in a community standard repository.'' This is good advice, especially
-considering your data is most likely to be reused if it is available on
-a community site. Community repositories can also provide guidance for
-best practices. As an example, if you are archiving sequencing data, a
-repository such as those run by the National Center for Biotechnology
-Information (NCBI) (e.g.~GenBank) provides not only a location for data
-archival, but also encourages a set of practices related to consistent
-data formatting and the inclusion of appropriate metadata. However, data
-storage policies are highly variable between repositories {[}15{]}. A
-data management plan utilizing best practices across all stages of the
-data life cycle will facilitate transition from local storage to
-repository {[}16{]}. Similarly, having such a plan can facilitate
-transition from repository to repository if funding runs out or
-requirements change. Good storage practices are important even (or
-especially) in cases where data may not fit with an existing repository,
-where only derived data products (versus raw data) are suitable for
-archiving, or in the case where an existing repository may have lax
-standards.
-
-This manuscript describes 10 simple rules for digital data storage that
+practices for data storage often begin and end with this statement:
+``Deposit your data in a community standard repository.'' This is good
+advice, especially considering your data is most likely to be reused if
+it is available on a community site. Community repositories can also
+provide guidance for best practices. As an example, if you are archiving
+sequencing data, a repository such as those run by the National Center
+for Biotechnology Information (NCBI) (e.g., GenBank) not only provides a
+location for data archival, but also encourages a set of practices
+related to consistent data formatting and the inclusion of appropriate
+metadata. However, data storage policies are highly variable between
+repositories {[}15{]}. A data management plan utilizing best practices
+across all stages of the data life cycle will facilitate transition from
+local storage to repository {[}16{]}. Similarly, having such a plan can
+facilitate transition from repository to repository if funding runs out
+or requirements change. Good storage practices are important even (or
+especially) in cases when data may not fit with an existing repository,
+when only derived data products (versus raw data) are suitable for
+archiving, or when an existing repository may have lax standards.
+
+This article describes ten simple rules for digital data storage that
 grew out of a long discussion among instructors for the Software and
@@
 
-\section*{Rule 1: Anticipate how your data will be
-used}\label{rule-1-anticipate-how-your-data-will-be-used}
-\addcontentsline{toc}{section}{Rule 1: Anticipate how your data will be
-used}
+\section*{Rule 1: Anticipate How Your Data Will Be
+Used}\label{rule-1-anticipate-how-your-data-will-be-used}
+\addcontentsline{toc}{section}{Rule 1: Anticipate How Your Data Will Be
+Used}
 
@@
 management, and release of data by having a clear roadmap of what to
-expect \emph{before} data acquisition starts. For instance:
+expect before data acquisition starts. For instance:
 
@@
   How will the raw data be received? Are they delivered by a machine or
-  software, or typed-in?
+  software, or typed in?
 \item
@@
 the data, both during and after the analysis. The required strategy will
-differ for datasets of varying size. Smaller datasets (e.g.~a few
+differ for datasets of varying size. Smaller datasets (e.g., a few
 megabytes in size) can be managed locally with a simple data management
-plan, whereas larger datasets (e.g.~gigabytes to petabytes) will in
+plan, whereas larger datasets (e.g., gigabytes to petabytes) will in
 almost all cases require careful planning and preparation (Rule 10).
@@
 
-\section*{Rule 2: Know your use case}\label{rule-2-know-your-use-case}
-\addcontentsline{toc}{section}{Rule 2: Know your use case}
+\section*{Rule 2: Know Your Use Case}\label{rule-2-know-your-use-case}
+\addcontentsline{toc}{section}{Rule 2: Know Your Use Case}
 
@@
   Are there restrictions or privacy concerns associated with the data
-  (e.g.~for survey results with personally identifiable information
-  (PII), threatened species, or confidential business information)?
+  (e.g., survey results with personally identifiable information
+  {[}PII{]}, threatened species, or confidential business information)?
 \item
@@
 questions to ask before starting data acquisition. But knowing the what,
-when, and how of \emph{your} use of the data will bring you close to a
-reliable roadmap on how to handle data from acquisition through
-publication and archival.
-
-\section*{Rule 3: Keep raw data raw}\label{rule-3-keep-raw-data-raw}
-\addcontentsline{toc}{section}{Rule 3: Keep raw data raw}
+when, and how of your use of the data will bring you close to a reliable
+roadmap on how to handle data from acquisition through publication and
+archival.
+
+\section*{Rule 3: Keep Raw Data Raw}\label{rule-3-keep-raw-data-raw}
+\addcontentsline{toc}{section}{Rule 3: Keep Raw Data Raw}
 
 Since analytical and data processing procedures improve or otherwise
-change over time, having access to the `raw' (unprocessed) data can
+change over time, having access to the ``raw'' (unprocessed) data can
 facilitate future re-analysis and analytical reproducibility. As
@@
 
-\section*{Rule 4: Store data in open
-formats}\label{rule-4-store-data-in-open-formats}
-\addcontentsline{toc}{section}{Rule 4: Store data in open formats}
+\section*{Rule 4: Store Data in Open
+Formats}\label{rule-4-store-data-in-open-formats}
+\addcontentsline{toc}{section}{Rule 4: Store Data in Open Formats}
 
 To maximize accessibility and long-term value, it is preferable to store
-data in formats whose specifications are freely available. The
-appropriate file type will depend on the data being stored (e.g.~numeric
-measurements, text, images, video), but the key idea is that accessing
-data should not require proprietary software, hardware, or purchase of a
-commercial license. Proprietary formats change, maintaining
-organizations go out of business, and changes in license fees make
-access to data in proprietary formats unaffordable and risky for
-end-users. Examples of open data formats include comma-separated values
-(CSV) for tabular data, hierarchical data format (HDF) {[}22{]} and
-NetCDF {[}23{]} for hierarchically structured scientific data, portable
-network graphics (PNG) for images, KML (or other Open Geospatial
-Consortium (OGC) format) for spatial data, and extensible markup
-language (XML) for documents. Examples of closed formats include DWG for
-AutoCAD drawings, Photoshop document (PSD) for bitmap images, Windows
-Media Audio (WMA) for audio recording files, and Microsoft Excel (XLS)
-for tabular data. Even if day-to-day processing uses closed formats
-(e.g., due to software requirements), data being stored for archival
-purposes should be stored in open formats. This is generally not
-prohibitive; most closed-source software products enable users to export
-data to an open format.
+data in formats that have freely available specifications. The
+appropriate file type will depend on the data being stored (e.g.,
+numeric measurements, text, images, video), but the key idea is that
+accessing data should not require proprietary software, hardware, or
+purchase of a commercial license. Proprietary formats change,
+maintaining organizations go out of business, and changes in license
+fees make access to data in proprietary formats unaffordable and risky
+for end-users. Examples of open data formats include comma-separated
+values (CSV) for tabular data, hierarchical data format (HDF) {[}22{]}
+and NetCDF {[}23{]} for hierarchically structured scientific data,
+portable network graphics (PNG) for images, KML (or other Open
+Geospatial Consortium {[}OGC{]} format) for spatial data, and extensible
+markup language (XML) for documents. Examples of closed formats include
+DWG for AutoCAD drawings, Photoshop document (PSD) for bitmap images,
+Windows Media Audio (WMA) for audio recording files, and Microsoft Excel
+(XLS) for tabular data. Even if day-to-day processing uses closed
+formats (e.g., due to software requirements), data being stored for
+archival purposes should be stored in open formats. This is generally
+not prohibitive; most closed-source software products enable users to
+export data to an open format.
 
@@
 stored in a format that computers can easily use for processing. This is
-especially crucial as datasets become larger. Making data easily-usable
+especially crucial as datasets become larger. Making data easily usable
 is best achieved by using standard data formats that have open
@@
 introduced during conversion between formats. Examples of
-machine-readable open formats that would \emph{not} be easy to process
-include data included in the text of a PDF file, or scanned images of
-tabular data from a paper source.
-
-\section*{Rule 5: Data should be structured for
-analysis}\label{rule-5-data-should-be-structured-for-analysis}
-\addcontentsline{toc}{section}{Rule 5: Data should be structured for
-analysis}
+machine-readable open formats that would not be easy to process include
+data included in the text of a PDF file, or scanned images of tabular
+data from a paper source.
+
+\section*{Rule 5: Data Should Be Structured for
+Analysis}\label{rule-5-data-should-be-structured-for-analysis}
+\addcontentsline{toc}{section}{Rule 5: Data Should Be Structured for
+Analysis}
 
@@
 a row, and each type of observational unit as a table (Fig. 1). The
-technical term for this structure is `Codd's 3rd normal form', but it
-has been made more accessible as theconcept of \emph{tidy} data
-{[}24{]}. When data is organized in this way, the duplication of
-information is reduced and it is easier to subset or summarize the
-dataset to include the variables or observations of interest.
+technical term for this structure is ``Codd's 3rd normal form'', but it
+has been made more accessible as the concept of tidy data {[}24{]}. When
+data is organized in this way, the duplication of information is reduced
+and it is easier to subset or summarize the dataset to include the
+variables or observations of interest.
 
@@
 ``write code for humans, write data for computers'' {[}25{]}. When data
-can be easily imported and manipulated using familiar software, whether
+can be easily imported and manipulated using familiar software (whether
 via a scripting language, a spreadsheet, or any other computer program
-that can import these common files, data becomes easier to re-use.
+that can import these common files), data becomes easier to re-use.
 Furthermore, having the source code for the software doing the analysis
@@
 
-\section*{Rule 6: Data should be uniquely
-identifiable}\label{rule-6-data-should-be-uniquely-identifiable}
-\addcontentsline{toc}{section}{Rule 6: Data should be uniquely
-identifiable}
+\section*{Rule 6: Data Should Be Uniquely
+Identifiable}\label{rule-6-data-should-be-uniquely-identifiable}
+\addcontentsline{toc}{section}{Rule 6: Data Should Be Uniquely
+Identifiable}
 
@@
 \href{http://www.dataone.org}{DataOne}, are able to provide these.
-Institutional initiatives also exist, and are known to your local
-librarians. Some repositories may require specific identifiers and these
-could change with time. For instance, NCBI sequence data will in the
-future only be identified by ``accession.version'' IDs. The ``GI''
+Institutional initiatives also exist and are known to your local
+librarians. Some repositories may require specific identifiers, and
+these could change with time. For instance, NCBI sequence data will in
+the future only be identified by ``accession.version'' IDs. The ``GI''
 identifiers (in use since 1994) will be retired in late 2016 {[}26{]}.
@@
 ambiguities: it mandates the date format \texttt{YYYY-MM-DD} (i.e.~from
-largest time unit to smallest). For example, the date ``February 1st,
-2015'', while written as \texttt{01-02-2015} in the UK and
+largest time unit to smallest). For example, the date ``February 1,
+2015,'' while written as \texttt{01-02-2015} in the UK and
 \texttt{02-01-2015} in the US, is unambiguous (\texttt{2015-02-01})
@@
 {[}27{]}. The CellPack datasets are an example of this {[}28{]}. A
-semantic version number takes the form: \texttt{Major.Minor.Patch}, e.g.
+semantic version number takes the form: \texttt{Major.Minor.Patch} e.g.,
 \texttt{0.2.7}. The \emph{major version} numbers should be incremented
@@
 \texttt{2.0.0} without changes to the data analysis. The \emph{minor
-version} should be bumped when a change has been made which is
-compatible with older versions of the data with the same major version.
-This means that any analysis that can be performed on version
-\texttt{1.0.0} of the data is repeatable with version \texttt{1.1.0} of
-the data. For example, adding a new year in a temporal survey will
-result in a bump in the minor version. The \emph{patch version} number
-is bumped when typos or bugs have been fixed. For example version
-\texttt{1.0.1} of a dataset may fix a typo in version \texttt{1.0.0}.
-
-\section*{Rule 7: Link relevant
-metadata}\label{rule-7-link-relevant-metadata}
-\addcontentsline{toc}{section}{Rule 7: Link relevant metadata}
+version} should be bumped when a change has been made that is compatible
+with older versions of the data with the same major version. This means
+that any analysis that can be performed on version \texttt{1.0.0} of the
+data is repeatable with version \texttt{1.1.0} of the data. For example,
+adding a new year in a temporal survey will result in a bump in the
+minor version. The \emph{patch version} number is bumped when typos or
+bugs have been fixed. For example version \texttt{1.0.1} of a dataset
+may fix a typo in version \texttt{1.0.0}.
+
+\section*{Rule 7: Link Relevant
+Metadata}\label{rule-7-link-relevant-metadata}
+\addcontentsline{toc}{section}{Rule 7: Link Relevant Metadata}
 
@@
 way to do this depends on the format of the data. Text files can contain
-meta-data in well-defined text files such as XML or JSON. Some file
-formats are self-documenting; for example NetCDF, HDF5, and many image
+metadata in well-defined text files such as XML or JSON. Some file
+formats are self-documenting; for example, NetCDF, HDF5, and many image
 file formats allow for embedded metadata {[}22,23{]}. In a relational
@@
 Ideally, a schema will be provided that also shows the linkages between
-data tables and metadata tables. Another--simpler--scenario is a set of
-flat (non-hierarchical) text files--in this case a semantically
+data tables and metadata tables. Another---simpler---scenario is a set
+of flat (non-hierarchical) text files---in this case a semantically
 versioned, compressed archive should be created that includes metadata.
@@
 
-\section*{Rule 8: Adopt the proper privacy
-protocols}\label{rule-8-adopt-the-proper-privacy-protocols}
-\addcontentsline{toc}{section}{Rule 8: Adopt the proper privacy
-protocols}
-
-In datasets where privacy is important, be sure to have a plan in place
-to protect data confidentiality. You should consider the different data
-stakeholders when developing privacy protocols for your data storage.
-These stakeholders include funding agencies, human subjects or entities,
-collaborators, and yourself. Both the NSF and NIH have data sharing
-policies in their grant guidelines to prevent sharing personally
+\section*{Rule 8: Adopt the Proper Privacy
+Protocols}\label{rule-8-adopt-the-proper-privacy-protocols}
+\addcontentsline{toc}{section}{Rule 8: Adopt the Proper Privacy
+Protocols}
+
+In datasets for which privacy is important, be sure to have a plan in
+place to protect data confidentiality. You should consider the different
+data stakeholders when developing privacy protocols for your data
+storage. These stakeholders include funding agencies, human subjects or
+entities, collaborators, and yourself. Both the National Science
+Foundation (NSF) and National Institutes of Health (NIH) have data
+sharing policies in their grant guidelines to prevent sharing personally
 identifiable information, and to anonymize data on human subjects.
@@
 In small datasets, a lookup table (protecting PII by removing it and
-replacing it with a unique id that maps to the sensitive data in an
+replacing it with a unique ID that maps to the sensitive data in an
 external dataset) is enough to anonymize a minimal amount of personal
@@
 
-\section*{Rule 9: Have a systematic backup
-scheme}\label{rule-9-have-a-systematic-backup-scheme}
-\addcontentsline{toc}{section}{Rule 9: Have a systematic backup scheme}
+\section*{Rule 9: Have a Systematic Backup
+Scheme}\label{rule-9-have-a-systematic-backup-scheme}
+\addcontentsline{toc}{section}{Rule 9: Have a Systematic Backup Scheme}
 
 Every storage medium can fail, and every failure can result in loss of
-data. Researchers should therefore back data up at all stages of the
+data. Researchers should therefore back up data at all stages of the
 research process. Data stored on local computers or institutional
@@
 used. Ideally you should have two on-site copies (such as on a computer,
-an external hard drive, or a tape) and one off-site copy (e.g.~cloud
+an external hard drive, or a tape) and one off-site copy (e.g., cloud
 storage){[}35{]}, with care taken to ensure that the off-site copy is as
@@
 
-\section*{Rule 10: The location and method of data storage depend on how
-much data you
-have}\label{rule-10-the-location-and-method-of-data-storage-depend-on-how-much-data-you-have}
-\addcontentsline{toc}{section}{Rule 10: The location and method of data
-storage depend on how much data you have}
+\section*{Rule 10: The Location and Method of Data Storage Depend on How
+Much Data You
+Have}\label{rule-10-the-location-and-method-of-data-storage-depend-on-how-much-data-you-have}
+\addcontentsline{toc}{section}{Rule 10: The Location and Method of Data
+Storage Depend on How Much Data You Have}
 
@@
 by environmental sensors, satellites, automated analytical tools,
-simulation models, and nucleic acid sequencers. Even larger data
-generating machines, like the Large Hadron Collider (LHC) and the Large
-Scale Synoptic Survey Telescope (LSST), generate many terabytes (TB) per
-day, rapidly accumulating to petabyte (PB) scale over the course of any
+simulation models, and nucleic acid sequencers. Even larger
+data-generating machines, like the Large Hadron Collider (LHC) and the
+Large Scale Synoptic Survey Telescope (LSST), generate many TB per day,
+rapidly accumulating to petabyte (PB) scale over the course of any
 particular study. While the cost of storage continues to decrease, the
@@
 this is less expensive, but can take longer to retrieve. Some storage
-systems automatically migrate `stale' files to longer term storage.
-Alternatively, some computing can be done `in the database' or `on disk'
-via database query languages (e.g.~SQL, MapReduce) that perform basic
-arithmetic, or via the use of procedural languages (e.g.~R, Python, C)
-embedded in the database server. Modern database technologies such as
-HDFS and Spark allow these computations to be done on data of almost any
-size. When data is larger than locally available RAM, it can be handled
-by conducting analyses on a `big memory' node, which most
+systems automatically migrate ``stale'' files to longer-term storage.
+Alternatively, some computing can be done ``in the database'' or ``on
+disk'' via database query languages (e.g., SQL, MapReduce) that perform
+basic arithmetic, or via the use of procedural languages (e.g., R,
+Python, C) embedded in the database server. Modern database technologies
+such as HDFS and Spark allow these computations to be done on data of
+almost any size. When data is larger than locally available RAM, it can
+be handled by conducting analyses on a ``big memory'' node, which most
 high-performance computing centers have deployed. Relying on tight
 software/hardware integration, these can allow for the analysis of
-datasets around 1-4 TB in size. This allows the user to read in and use
+datasets around 1--4 TB in size. This allows the user to read in and use
 a large dataset without special tools.
@@
 If you regularly only need access to a small subset of your data or need
-to share it with many collaborators, a web-based API (Application
+to share it with many collaborators, an online API (Application
 Programming Interface) might be a good solution. Using this method, many
-users can send requests to a web service which can subset the data,
+users can send requests to a online service that can subset the data,
 perform in-database computation, and return smaller volumes of data as
-specific slices. Tools based on web services make it easier to find and
-download data, and facilitate analysis via reproducible scripts.
-However, they can also lead to excessive and careless abuse of resources
-without proper safeguards in place. The time required to re-download and
-recompute results can be reduced by `caching'. Caching stores copies of
-downloads and generated files that are recognized when the same script
-is run multiple times.
+specific slices. Tools based on online services make it easier to find
+and download data, and they facilitate analysis via reproducible
+scripts. However, they can also lead to excessive and careless abuse of
+resources without proper safeguards in place. The time required to
+re-download and re-compute results can be reduced by ``caching''.
+Caching stores copies of downloads and generated files that are
+recognized when the same script is run multiple times.
 
@@
 
-\section*{Glossary used in the
-manuscript}\label{glossary-used-in-the-manuscript}
-\addcontentsline{toc}{section}{Glossary used in the manuscript}
+\section*{Glossary}\label{glossary}
+\addcontentsline{toc}{section}{Glossary}
 
@@
   \textbf{Comma-Separated Values} (CSV) and \textbf{Tab-Separated
-  Values} (TSV) are plain text file formats used to store tabular data
-  where each row is represented by a line in the file, and each field
-  (column) is separated by a comma (for CSV) or by the Tab character
+  Values} (TSV) are plain text file formats used to store tabular data,
+  in which each row is represented by a line in the file, and each field
+  (column) is separated by a comma (for CSV) or by the tab character
   (for TSV).
@@
   \textbf{FASTA} is a simple and widely used file format used to
-  represent sequences of nucleotides or amino-acids in plain text making
-  it easy to manipulate programmatically.
+  represent sequences of nucleotides or amino acids in plain text,
+  making it easy to manipulate these programmatically.
 \item
@@
   to how a hard drive is organized with directories and files. It is
-  maintained by the non-profit HDF Group, a spin off of the National
+  maintained by the non-profit HDF Group, a spin-off of the National
   Center for Supercomputing Applications (NCSA).
@@
   typically used to store arbitrarily structured data in the form of
-  keys and values. It can be used to store non-relational databases as
+  keys and values. It can be used to store non-relational databases, as
   it does not rely on a tabular data format. In many respects, it has
@@
   file format designed to store large datasets in array-oriented
-  scientific data as typically used in the geosciences. It is maintained
+  scientific data, typically used in the geosciences. It is maintained
   by Unidata, a non-profit member of the University Corporation for
-  Atmospheric Research (UCAR) which is funded by the National Science
+  Atmospheric Research (UCAR), which is funded by the National Science
   Foundation.
@@
   \textbf{Hadoop Distributed File System} (HDFS) is a Java-based file
-  system where data is stored in small chunks across multiple redundant
-  nodes.
+  system in which data is stored in small chunks across multiple
+  redundant nodes.
 \item
@@
   large datasets in parallel computing environments. Such programs are
-  composed of a \textbf{map} procedure where the dataset is sliced into
-  several pieces, and a \textbf{reduce} procedure where summary
+  composed of a \textbf{map} procedure in which the dataset is sliced
+  into several pieces, and a \textbf{reduce} procedure in which summary
   operations are then applied to each of the slices.
@@
   SHA algorithms are often used in preference to similar tools such as
-  MD5 (mentioned in Rule 8), which are no longer secure. All hashing
-  algorithms are vulnerable to brute force attacks. Key Derivation
-  Function (KDF) implementations like BCrypt and PBKDF2 are considered
-  significantly more secure, but by design more costly to compute.
-\item
-  Apache \textbf{Spark} is an open source computing platform for
-  querying large data sets in memory, in contrast to on-disk based
+  MD5 (mentioned in Rule 3 and 8), which are no longer secure. All
+  hashing algorithms are vulnerable to brute force attacks. Key
+  Derivation Function (KDF) implementations like BCrypt and PBKDF2 are
+  considered significantly more secure, but by design more costly to
+  compute.
+\item
+  Apache \textbf{Spark} is an open-source computing platform for
+  querying large datasets in memory, in contrast to on--disk--based
   methods like MapReduce.
@@
   needed to store a single character of text in a computer. The prefixes
-  mega-, giga-, tera- and peta- refer to the international system of
+  mega-, giga-, tera-, and peta- refer to the international system of
   units for the multiple of the unit, and correspond to \(10^{6}\),
-  \(10^{9}\), \(10^{12}\), and \(10^{15}\) abbreviated M, G, T and P
+  \(10^{9}\), \(10^{12}\), and \(10^{15}\), abbreviated M, G, T, and P,
   respectively.
@@
   \textbf{Digital Object Identifier} (DOI) provides unique and
-  persistent identifiers for electronic documents (in particular journal
-  articles and datasets) on the internet. The uniqueness of the
+  persistent identifiers for electronic documents (in particular,
+  journal articles and datasets) on the internet. The uniqueness of the
   identifiers is guaranteed by a central registry. By dissociating the
@@
 \item
-  \textbf{Persistent Uniform Resource Locator} (PURL) is an URL used to
+  \textbf{Persistent Uniform Resource Locator} (PURL) is a URL used to
   redirect to the location of an electronic object on the internet. DOI
@@
 We thank K. Hinsen, G. Pugh, and W. Turner for comments on the draft
-manuscript. Contributions from JWH, an employee of the U.S.
+manuscript. Contributions from J. W. Hollister, an employee of the US
 Environmental Protection Agency (EPA) have not been subjected to the
@@
 the views of the Agency. Mention of trade names or commercial products
-does not constitute endorsement or recommendation for use. SM would like
-to thank Dr.~Laurence Tratt for reviewing an early draft of this paper.
+does not constitute endorsement or recommendation for use. S. Mount
+would like to thank Dr.~Laurence Tratt for reviewing an early draft of
+this paper.
 
